### 2017-05-16 paper/paper.tex
Adding the paper files

% Template for ICASSP-2010 paper; to be used with:
%          mlspconf.sty  - ICASSP/ICIP LaTeX style file adapted for MLSP, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{amsmath,graphicx,mlspconf}

%Select one of the four copyright notices below. Only required for the camera paper submission

%For papers in which all authors are employed by the US government, the copyright notice is:
\copyrightnotice{U.S.\ Government work not protected by U.S.\ copyright}

%For papers in which all authors are employed by a Crown government (UK, Canada, and Australia), the copyright notice is:
\copyrightnotice{978-1-5090-6341-3/17/\$31.00 {\copyright}2017 Crown}

%For papers in which all authors are employed by the European Union, the copyright notice is:
\copyrightnotice{978-1-5090-6341-3/17/\$31.00 {\copyright}2017 European Union}

%For all other papers the copyright notice is:
\copyrightnotice{978-1-5090-6341-3/17/\$31.00 {\copyright}2017 IEEE}

\toappear{2017 IEEE International Workshop on Machine Learning for Signal Processing, Sept.\ 25--28, 2017, Tokyo, Japan}

% Example definitions.
% --------------------
\def\x{{\mathbf x}}
\def\L{{\cal L}}

% Title.
% ------
\title{AUTHOR GUIDELINES FOR MLSP PROCEEDINGS MANUSCRIPTS}
%
% Single address.
% ---------------
\name{Author(s) Name(s) omitted for double blind review\thanks{Thanks to XYZ agency for funding.}}
\address{Author Affiliation(s) omitted for double blind review}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%   Email A-B
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D
%   Email C-D}
%
\begin{document}
%\ninept
%

\maketitle
%
\begin{abstract}
The abstract should appear at the top of the left-hand column of text, about
0.5 inch (12 mm) below the title area and no more than 3.125 inches (80 mm) in
length.  Leave a 0.5 inch (12 mm) space between the end of the abstract and the
beginning of the main text.  The abstract should contain about 100 to 150
words, and should be identical to the abstract text submitted electronically
along with the paper cover sheet.  All manuscripts must be in English, printed
in black ink.
\end{abstract}
%
\begin{keywords}
One, two, three, four, five
\end{keywords}
%
\section{Introduction}
\label{sec:intro}

This is a good paper.

\section{CNN-CNN Autoencoder} 
The approximation $\hat X$ for a given spectrogram $X$ is computed as follows:

\begin{align}
	\hat H(k,t) =& ( \sum_{f,t'} X(f,t-t') F_e(f,t',k) ) \notag \\
	\hat X(f,t) =& \sum_k \sum_{t'} \hat H(k,t-t') F_d(f, t',k)
\end{align}

\section{RNN-CNN Autoencoder}
This is the same as CNN-CNN case, except the computation of the activations $H$. In the CNN encoder, each filter was of finite length. With RNN-CNN version, we are attempting to use an infinite length filter. The computation of $\hat H$ is as follows:

\begin{align}
	Z(:,k,t) =& \sigma( W Z(:,k,t-1) + U X(:,t)) \notag \\
	\hat H(k,t) =& \sum_f Z(f,k,t)
\end{align}

\section{REFERENCES}
\label{sec:ref}

List and number all bibliographical references at the end of the paper.  The references can be numbered in alphabetic order or in order of appearance in the document.  When referring to them in the text, type the corresponding reference number in square brackets as shown at the end of this sentence \cite{C2}.

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\bibliographystyle{IEEEbib}
\bibliography{strings,refs}

\end{document}

### 2017-05-18 main.py
I have just finished the sourceseparation infrastructure

@@ -8,6 +8,7 @@ import multiprocessing
 from multiprocessing import Process
 from multiprocessing.pool import Pool
 from rnns import *
+from music_utilities import *
 
 def model_driver(d,data):
     """This function builds the computation graph for the specified model
@@ -19,24 +20,61 @@ def model_driver(d,data):
     tf.reset_default_graph()
    
     #build the first graph
-    rnn1 = rnn( model_specs = d, initializer = d['init']) 
-    rnn1_handles = rnn1.build_graph()  
-
     config = tf.ConfigProto(log_device_placement = False)
     config.gpu_options.allow_growth=True
     config.allow_soft_placement=True
 
+    #build graphs
+    with tf.variable_scope("model1"):
+        rnn1 = rnn( model_specs = d, initializer = d['init'], model_num = 1)
+        rnn1_handles = rnn1.build_graph()  
+    
     with tf.Session(config = config) as sess:
         sess.run(tf.initialize_all_variables())
+        all_times, tr_logls, _, _ = rnn1.optimizer(data = data['Train1'], 
+                rnn_handles = rnn1_handles, sess = sess)
+        
+        vars_np1 = rnn1.save_modelvars_np(sess)
 
-        all_times, tr_logls, test_logls, valid_logls = rnn1.optimizer(data = data, rnn_handles = rnn1_handles, sess = sess) 
+    tf.reset_default_graph()
 
-    max_valid = np.max(np.array(valid_logls)) #kind of unnecessary
-    max_test = np.max( np.array(test_logls))
+    with tf.variable_scope("model2"):
+        rnn2 = rnn( model_specs = d, initializer = d['init'], model_num = 2)
+        rnn2_handles = rnn2.build_graph()  
+
+    with tf.Session(config = config) as sess:
+        sess.run(tf.initialize_all_variables())
+        all_times, tr_logls, _, _ = rnn2.optimizer(data = data['Train2'], 
+                rnn_handles = rnn2_handles, sess = sess) 
+
+        vars_np2 = rnn2.save_modelvars_np(sess)
+
+    tf.reset_default_graph()
+
+    all_vars_np = vars_np1 + vars_np2
+
+    rnn_sep_opt = rnn( model_specs = d, initializer = all_vars_np, input_opt = True)
+    rnn_sep_opt_handles = rnn_sep_opt.build_separation_graph()
+
+    with tf.Session(config = config) as sess:
[…]

### 2017-05-18 music_utilities.py

@@ -1,0 +1,182 @@
+from numpy import *
+import scipy.io.wavfile as wavfile
+import os
+eps = finfo( float32).eps;
+
+def tset( mf = None, ff = None, dr = None):
+    # Where the files are
+    p = '/usr/local/timit/timit-wav/train/'
+
+    # Pick a speaker directory
+    if dr is None:
+        dr = random.randint( 1, 8)
+    p += 'dr%d/' % dr
+
+    # Get two random speakers
+    if mf is None:
+        mf = [name for name in os.listdir( p) if name[0] == 'm']
+        mf = random.choice( mf)
+    if ff is None:
+        ff = [name for name in os.listdir( p) if name[0] == 'f']
+        ff = random.choice( ff)
+    print ('dr%d/' % dr, mf, ff)
+
+    # Load all the wav files
+    ms = [wavfile.read(p+mf+'/'+n)[1] for n in os.listdir( p+mf) if 'wav' in n]
+    fs = [wavfile.read(p+ff+'/'+n)[1] for n in os.listdir( p+ff) if 'wav' in n]
+
+    # Find suitable test file pair
+    l1 = list( map( lambda x : x.shape[0], ms))
+    l2 = list( map( lambda x : x.shape[0], fs))
+
+    d = array( [[abs(t1-t2) for t1 in l1] for t2 in l2])
+    i = argmin( d)
+    l = max( [l1[i%10], l2[int(i/10)]])
+    ts = [pad( ms[i%10], (0,l-l1[i%10]), 'constant'), pad( fs[int(i/10)], (0,l-l2[int(i/10)]), 'constant')]
+
+    # Get training data
+    ms.pop( i%10)
+    fs.pop( int(i/10))
+    tr = [concatenate(ms), concatenate(fs)]
+
+    return list(map( lambda x : (x-mean(x))/std(x), ts)), list(map( lambda x : (x-mean(x))/std(x), tr))
+
+
+def sound_set( tp):
+    import scipy.io.wavfile as wavfile
+
+    # Two sinusoids signal
+    if tp == 1:
+        l = 8*1024
+        def clip0( x):
+            return x * (x>0)
+        z1 = clip0( sin( 3*linspace( 0, 2*pi, l))) * sin( 1099*linspace( 0, 2*pi, l))
+        z2 = clip0( sin( 2*linspace( 0, 2*pi, l))) * sin( 3222*linspace( 0, 2*pi, l))
[…]

### 2017-05-18 rnns.py

@@ -10,20 +10,131 @@ import itertools
 import sys
 import os
 import matplotlib.pyplot as plt
+import scipy.io.wavfile as wavfile
+from music_utilities import *
+
+
+class ModLSTMCell(tf.contrib.rnn.RNNCell):
+    """Modified LSTM Cell """
+
+    def __init__(self, num_units, initializer = tf.contrib.layers.xavier_initializer(uniform=True, seed=2, dtype=tf.float32), wform = 'diagonal', input_opt = False, model_num = None):
+        self._num_units = num_units
+        self.init = initializer
+        self.wform = wform 
+        self.input_opt = input_opt
+        self.model_num = model_num
+
+    @property
+    def state_size(self):
+        return tf.contrib.rnn.LSTMStateTuple(self._num_units, self._num_units)
+
+    @property
+    def output_size(self):
+        return self._num_units
+
+    def __call__(self, inputs, state, scope=None):
+        with tf.variable_scope(scope or type(self).__name__):  # "GRUCell"
+            
+            c, h = state
+            init = self.init
+            self.L1 = inputs.get_shape().as_list()[1]
+           
+            mats, biases = self.get_params_parallel()
+            if self.wform == 'full' or self.wform == 'diag_to_full':
+                
+                res = tf.matmul(tf.concat([h,inputs],axis=1),mats)
+                res_wbiases = tf.nn.bias_add(res, biases)
+           
+                i,j,f,o = tf.split(res_wbiases,num_or_size_splits=4,axis=1) 
+            elif self.wform == 'diagonal':
+                h_concat = tf.concat([h,h,h,h],axis=1)
+
+                W_res = tf.multiply(h_concat,mats[0])
+
+                U_res = tf.matmul(inputs,mats[1])
+
+                res = tf.add(W_res,U_res)
+                res_wbiases = tf.nn.bias_add(res, biases)
+
+                i,j,f,o = tf.split(res_wbiases,num_or_size_splits=4,axis=1) 
+                         
+            new_c = (c * tf.nn.sigmoid(f) + tf.nn.sigmoid(i)*tf.nn.tanh(j))
+
+            new_h = tf.nn.tanh(new_c) * tf.nn.sigmoid(o)
+            new_state = tf.contrib.rnn.LSTMStateTuple(new_c, new_h)
+        return new_h, new_state
+
+
+    def get_params_parallel(self):
+        if self.input_opt:
+            if self.wform == 'full':
+
[…]

### 2017-05-19 main.py
I have fixed the error in separation, and added activation function selection mechanism

@@ -19,7 +19,6 @@ def model_driver(d,data):
     # Reset graph
     tf.reset_default_graph()
    
-    #build the first graph
     config = tf.ConfigProto(log_device_placement = False)
     config.gpu_options.allow_growth=True
     config.allow_soft_placement=True
@@ -59,7 +58,6 @@ def model_driver(d,data):
     with tf.Session(config = config) as sess:
         sess.run(tf.initialize_all_variables())
     
-        pdb.set_trace()
         x1hat, x2hat = rnn_sep_opt.input_optimizer(data = data['Test'], 
                                     rnn_handles = rnn_sep_opt_handles, sess = sess) 
 
@@ -71,6 +69,7 @@ def model_driver(d,data):
     o2 = d['FE'].ife( x2hat * (mixture/(x1hat+x2hat)), phase)
     sxr = bss_eval( o1, 0, vstack( (Z[2],Z[3]))) + bss_eval( o2, 1, vstack( (Z[2],Z[3])))
     
+    print('BSS eval values are: ',str(sxr))
     pdb.set_trace()
 
     res_dictionary = {'valid':  np.array(valid_logls), 
@@ -177,6 +176,8 @@ input_dictionary = {'seedin' : [1144, 1521], #setting the random seed. First is 
             'lr_min':-4, 'lr_max':-2, #the lower and upper limits for the exponent of the learning rate
             'num_layers_min':1, 'num_layers_max':3, #lower and upper limits for number of layers
             'optimizer':'RMSProp', #options are, Adam, RMSProp, Adadelta
-            'notes':'I am trying the gated_wf model; with n layers in {1,3} on ptb. K is in range {200,...,400}, I have just added iterations over batches'}
+            'activation':'softplus',
+            'separation_method':'complete',
+            'notes':''} 
 
 perfs = main(input_dictionary)

### 2017-05-19 rnns.py

@@ -133,6 +133,12 @@ class rnn(object):
         'this function builds a graph with the specifications in self.model_specs'
         
         d = self.model_specs #unpack the model specifications
+        
+        if d['activation'] == 'relu':
+            act = tf.nn.relu
+        elif d['activation'] == 'softplus':
+            act = tf.nn.softplus
+
         with tf.device('/' + d['device']):
             x = tf.placeholder(tf.float32, [None, d['batchsize'+str(self.model_num)], d['L1']],"x")
             mask = tf.placeholder(tf.float32, [None])
@@ -142,6 +148,7 @@ class rnn(object):
             seq_lens = tf.placeholder(tf.int32, [None])
             
             hhat = self.define_model(x, seqlens = seq_lens, dropout_kps = dropout_kps)
+            hhat = act(hhat) 
 
             with tf.variable_scope("decoder"):
                     V_initializer = b_initializer = tf.contrib.layers.xavier_initializer(uniform=True, seed=2, dtype=tf.float32)
@@ -151,7 +158,7 @@ class rnn(object):
                     b = tf.get_variable("b", dtype= tf.float32, 
                             shape = [d['L2']], initializer = b_initializer)  
 
-            yhat = tf.nn.relu(tf.matmul(hhat,V) + tf.reshape(b, (1, d['L2'])))
+            yhat = act(tf.matmul(hhat,V) + tf.reshape(b, (1, d['L2'])))
 
             #compute the number of parameters to be trained
             tvars = tf.trainable_variables()
@@ -171,7 +178,8 @@ class rnn(object):
                 self.tvars_names = [var.name for var in tvars] 
 
             #define the cost         
-            cost = tf.reduce_sum(tf.square(y - yhat))
+            eps = 1e-16
+            cost = tf.reduce_mean( y*(tf.log( y + eps ) - tf.log( yhat + eps ))  - y + yhat ) 
 
             #define the optimizer
             if d['optimizer'] == 'Adam':
@@ -214,7 +222,7 @@ class rnn(object):
         # unpack model specifications 
         d = self.model_specs
         wform, model, K, num_layers, mapping_mode, L1, L2 = d['wform'], d['model'], d['K'], d['num_layers'], d['mapping_mode'], d['L1'], d['L2']
-
+        
         if model in onedir_rnns:
 
             if model == 'ss_lstm': 
@@ -257,39 +265,50 @@ class rnn(object):
[…]

### 2017-05-29 main.py
Implemented the convolutive encoder

@@ -172,13 +172,13 @@ wform = 'full'# either diagonal or full
 input_dictionary = {'seedin' : [1144, 1521], #setting the random seed. First is for numpy, second is for tensorflow 
             'task' : 'source_sep', #this helps us how to load the data with the load_data function in rnns.py 
             'data' : 'timit', #the dataset, options are inside the load_data function 
-            'encoder': 'mb_mod_lstm', #options are: mod_lstm (our custom cell), lstm (tensor flow's lstm cell), gated_wf (our custom gru cell), gru (tensor flow's gru cell), mod_rnn (our custom rnn cell) 
+            'encoder': 'convolutive', #options are: feed_forward, convolutive, mb_mod_lstm
             'decoder': 'convolutive',
             'wform' : wform, 
             'wform_global' : wform,
             'num_configs' : 60, #number of hyper parameter configurations to be tried 
             'start' : 0,  #this is used to start from a certain point (can be useful with fixed seed, or when hyper-parameters are loaded) 
-            'EP' : 2000, #number of epochs per run 
+            'EP' : 20, #number of epochs per run 
             'dropout' : [1, 1], #first is the input second is the output keep probability 
             'device' : 'cpu:0', #the device to be used in the computations 
             'server': socket.gethostname(),

### 2017-05-29 rnns.py

@@ -356,12 +356,12 @@ class rnn(object):
                         else:
                             b= tf.constant(var[1])
                 else:
-                    V_initializer = b_initializer = tf.contrib.layers.xavier_initializer(uniform=True, seed=2, dtype=tf.float32)
+                    initializer = tf.contrib.layers.xavier_initializer(uniform=True, seed=2, dtype=tf.float32)
                     
                     V = tf.get_variable("V", dtype= tf.float32, 
-                            shape = [d['L1'], d['K']], initializer = V_initializer)  
+                            shape = [d['L1'], d['K']], initializer = initializer)  
                     b = tf.get_variable("b", dtype= tf.float32, 
-                            shape = [d['K']], initializer = b_initializer)  
+                            shape = [d['K']], initializer = initializer)  
 
             x = tf.transpose(x, [1, 0, 2])
             x = tf.unstack(x, axis = 0)
@@ -370,6 +370,43 @@ class rnn(object):
             outputs = ( tf.matmul(x,V) + b ) 
 
             return outputs
+        
+        elif model == 'convolutive':
+            
+            with tf.variable_scope('encoder'):
+                if self.input_opt:
+                    vars_to_use = [var for var in self.initializer if 'model'+ str(def_model_num)+'/encoder' in var[0]] 
+                    for var in vars_to_use:
+                        if '/fltr' in var[0]:
+                            fltr = tf.constant(var[1])
+                        else:
+                            b = tf.constant(var[1])
+                else:
+                    initializer = tf.contrib.layers.xavier_initializer(uniform=True, seed=2, dtype=tf.float32)
+                    fltr = tf.get_variable("fltr", dtype= tf.float32, 
+                        shape = [d['ntaps'], 1, d['L1'], d['K']], 
+                        initializer = initializer)  
+                    b = tf.get_variable("b", dtype= tf.float32, 
+                        shape = [d['K']], initializer = initializer)  
+
+            x = tf.transpose(x, [1, 0, 2])
+            x = tf.unstack(x, axis = 0)
+            x = tf.concat(x, axis = 0)
+
+            x_rev = tf.reverse( x, axis = [0])
+            x_pad = tf.concat([x_rev, tf.zeros([d['ntaps']-1,d['L1']])],
+                                axis = 0)
[…]

### 2017-05-31 paper/paper.tex
A plan for the paper

@@
 \begin{abstract}
-The abstract should appear at the top of the left-hand column of text, about
-0.5 inch (12 mm) below the title area and no more than 3.125 inches (80 mm) in
-length.  Leave a 0.5 inch (12 mm) space between the end of the abstract and the
-beginning of the main text.  The abstract should contain about 100 to 150
-words, and should be identical to the abstract text submitted electronically
-along with the paper cover sheet.  All manuscripts must be in English, printed
-in black ink.
+   In this paper we propose a convolutive and recurrent neural network based extensions to autoencoders for source separation. 
 \end{abstract}
@@
 
-This is a good paper.
+Talk about what we are doing and why it is interesting. \\
 
-\section{CNN-CNN Autoencoder} 
+Give some background.
+
+\section{Autoencoders}
+
+\subsection{Feed-forward Autoencoder} 
+Define standard autoencoder. Provide a toy example to illustrate the shortcomings of this.
+
+\subsection{CNN-CNN Autoencoder} 
 The approximation $\hat X$ for a given spectrogram $X$ is computed as follows:
@@
 \begin{align}
-	\hat H(k,t) =& ( \sum_{f,t'} X(f,t-t') F_e(f,t',k) ) \notag \\
-	\hat X(f,t) =& \sum_k \sum_{t'} \hat H(k,t-t') F_d(f, t',k)
+	\hat H(k,t) =& \sigma_1\left ( \sum_{f,t'} X(f,t-t') F_e(f,t',k) \right ) \notag \\
+	\hat X(f,t) =& \sigma_2 \left ( \sum_k \sum_{t'} \hat H(k,t-t') F_d(f, t',k) \right )
 \end{align}
 
-\section{RNN-CNN Autoencoder}
+\subsection{Multilayer of CNN-CNN Autoencoder} 
+This loses the interpretability, but maybe good for accuracy?
+
+\subsection{RNN-CNN Autoencoder}
+
 This is the same as CNN-CNN case, except the computation of the activations $H$.
@@
 
+Give a toy example which shows what this model can do that CNN-CNN can not.
+
+\section{Experiments}
+
+Try each model in a given $K$ range in speech-speech source separation task.
+
 \section{REFERENCES}
 \label{sec:ref}
-
-List and number all bibliographical references at the end of the paper.
-The references can be numbered in alphabetic order or in order of appearance in the document.
-When referring to them in the text, type the corresponding reference number in square brackets as shown at the end of this sentence \cite{C2}.
 

### 2017-06-01 main.py
I have set up the experimental set-up which sweeps in a given range

@@ -172,18 +172,18 @@ wform = 'full'# either diagonal or full
 input_dictionary = {'seedin' : [1144, 1521], #setting the random seed. First is for numpy, second is for tensorflow 
             'task' : 'source_sep', #this helps us how to load the data with the load_data function in rnns.py 
             'data' : 'timit', #the dataset, options are inside the load_data function 
-            'encoder': 'convolutive', #options are: feed_forward, convolutive, mb_mod_lstm
-            'decoder': 'convolutive',
+            'encoder': 'mb_mod_lstm', #options are: feed_forward, convolutive, mb_mod_lstm
+            'decoder': 'feed_forward',
             'wform' : wform, 
             'wform_global' : wform,
             'num_configs' : 60, #number of hyper parameter configurations to be tried 
             'start' : 0,  #this is used to start from a certain point (can be useful with fixed seed, or when hyper-parameters are loaded) 
-            'EP' : 20, #number of epochs per run 
+            'EP' : 2000, #number of epochs per run 
             'dropout' : [1, 1], #first is the input second is the output keep probability 
-            'device' : 'cpu:0', #the device to be used in the computations 
+            'device' : 'gpu:1', #the device to be used in the computations 
             'server': socket.gethostname(),
             'verbose': True, #this prints out the batch location
-            'load_hparams': False, #this loads hyper-parameters from a results file
+            'load_hparams': True, #this loads hyper-parameters from a results file
             'count_mode': False, #if this is True, the code will stop after printing the number of trainable parameters
             'init':'xavier', #initialization method, options are 'xavier','random_unform' 
             'lr_min':-4, 'lr_max':-2, #the lower and upper limits for the exponent of the learning rate

### 2017-06-01 rnns.py

@@ -5,7 +5,7 @@ import pdb
 import time
 import pandas as pd
 import sklearn.feature_extraction.text as ft
-import _pickle as pickle
+#import _pickle as pickle
 import itertools
 import sys
 import os
@@ -635,23 +635,23 @@ def return_Klimits(model, wform, data):
 
     if model in ['mod_lstm', 'lstm']:
         min_params = 1e1; max_params =  7e7 # in our waspaa paper we basically did not use lower and upper bounds for number of parameters
-        K_min, K_max = 30, 350
+        K_min, K_max = 10, 100
 
     elif model in ['mb_mod_lstm']:
         min_params = 1e1; max_params = 7e7 
-        K_min, K_max = 10, 10
+        K_min, K_max = 10, 100
 
     elif model == 'feed_forward':
         min_params = 1e1; max_params = 7e7 
-        K_min, K_max = 10, 10
+        K_min, K_max = 10, 100
 
     elif model == 'convolutive':
         min_params = 1e1; max_params = 7e7
-        K_min, K_max = 10, 10
+        K_min, K_max = 10, 100
 
     return K_min, K_max, min_params, max_params 
 
-def generate_random_hyperparams(lr_min, lr_max, K_min, K_max, num_layers_min, num_layers_max,load_hparams):
+def generate_random_hyperparams(lr_min, lr_max, K_min, K_max, num_layers_min, num_layers_max,load_hparams, num_configs = 60):
     """This function generates random hyper-parameters for hyperparameter search"""
 
     #this is for new random parameters
@@ -665,14 +665,18 @@ def generate_random_hyperparams(lr_min, lr_max, K_min, K_max, num_layers_min, nu
 
     #this loads hyperparameters from an existing file
     else:
-        exp_data = np.load('experiment_data/nmf_data_timit_model_bi_mod_lstm_diag_to_full_device_cpu:0_1490813245.npy')[load_hparams[1]]
-        lr = exp_data['LR']
-        K = exp_data['K']
-        num_layers = exp_data['num_layers']
-        try:
-            momentum = exp_data['num_layers']
-        except:
-            momentum = None
+        #exp_data = np.load('experiment_data/nmf_data_timit_model_bi_mod_lstm_diag_to_full_device_cpu:0_1490813245.npy')[load_hparams[1]]
+        #pick K in this range
+        Krange = np.round(np.arange(start = K_min, stop = K_max+1, step = 10))
[…]

### 2017-06-01 rnns.py
I have fixed the error with multiple basis rnn encoder

@@ -338,7 +338,7 @@ class rnn(object):
             outputs = tf.unstack(outputs,axis = 0)
             outputs = tf.concat(outputs, axis = 0)
 
-            outputs_split = tf.split(outputs, axis = 1, num_or_size_splits = int(K))
+            outputs_split = tf.split(outputs, axis = 1, num_or_size_splits = int(d['K_in']))
             outputs_stack = tf.stack(outputs_split)
 
             outputs = tf.reduce_sum( outputs_stack, axis = 0) 
@@ -639,7 +639,7 @@ def return_Klimits(model, wform, data):
 
     elif model in ['mb_mod_lstm']:
         min_params = 1e1; max_params = 7e7 
-        K_min, K_max = 10, 100
+        K_min, K_max = 10, 30 
 
     elif model == 'feed_forward':
         min_params = 1e1; max_params = 7e7 

### 2017-06-02 main.py
working on dataset selection

@@ -90,11 +90,7 @@ def model_driver(d,data):
 
 def main(dictionary):
 
-    # first get the data and the resulting model parameters  
-    data, parameters = load_data(dictionary = dictionary)
-    dictionary.update(parameters) # add the necessary model parameters to the dictionary here
-    
-    ### next thing is determining the hyperparameters
+        ### next thing is determining the hyperparameters
     np.random.seed( dictionary['seedin'][0] )
     tf.set_random_seed( dictionary['seedin'][1] ) 
     timestamp = str(round(time.time()))
@@ -109,6 +105,11 @@ def main(dictionary):
 
     records = [] #information will accumulate in this
     for i in range(dictionary['num_configs']):
+        
+        # first get the data and the resulting model parameters  
+        data, parameters = load_data(dictionary = dictionary, ntrial = i)
+        dictionary.update(parameters) # add the necessary model parameters to the dictionary here
+
         while True:  
             try:
                 lr, K, num_layers, momentum = generate_random_hyperparams(
@@ -173,12 +174,12 @@ input_dictionary = {'seedin' : [1144, 1521], #setting the random seed. First is 
             'task' : 'source_sep', #this helps us how to load the data with the load_data function in rnns.py 
             'data' : 'timit', #the dataset, options are inside the load_data function 
             'encoder': 'mb_mod_lstm', #options are: feed_forward, convolutive, mb_mod_lstm
-            'decoder': 'feed_forward',
+            'decoder': 'convolutive',
             'wform' : wform, 
             'wform_global' : wform,
             'num_configs' : 60, #number of hyper parameter configurations to be tried 
             'start' : 0,  #this is used to start from a certain point (can be useful with fixed seed, or when hyper-parameters are loaded) 
-            'EP' : 2000, #number of epochs per run 
+            'EP' : 1, #number of epochs per run 
             'dropout' : [1, 1], #first is the input second is the output keep probability 
             'device' : 'gpu:1', #the device to be used in the computations 
             'server': socket.gethostname(),

### 2017-06-02 music_utilities.py

@@ -1,10 +1,13 @@
 from numpy import *
 import scipy.io.wavfile as wavfile
 import os
+import pdb 
+
 eps = finfo( float32).eps;
 
 def tset( mf = None, ff = None, dr = None):
     # Where the files are
+    # p = '~/Dropbox/timit/timit-wav/train/' 
     p = '/usr/local/timit/timit-wav/train/'
 
     # Pick a speaker directory
@@ -42,7 +45,7 @@ def tset( mf = None, ff = None, dr = None):
     return list(map( lambda x : (x-mean(x))/std(x), ts)), list(map( lambda x : (x-mean(x))/std(x), tr))
 
 
-def sound_set( tp):
+def sound_set( tp , selector):
     import scipy.io.wavfile as wavfile
 
     # Two sinusoids signal
@@ -79,7 +82,7 @@ def sound_set( tp):
 
     # TIMIT male/female set
     elif tp == 3:
-        ts,tr = tset( 'fbjl0', 'mwsh0', 5)
+        ts,tr = tset( 'fbjl0', 'mwsh0', selector )
         # ts,tr = tset( 'falr0', 'mtqc0', 4)
         # ts,tr = tset()
         sr = 16000

### 2017-06-02 rnns.py

@@ -680,7 +680,7 @@ def generate_random_hyperparams(lr_min, lr_max, K_min, K_max, num_layers_min, nu
 
     return lr, K, num_layers, momentum
 
-def load_data(dictionary):
+def load_data(dictionary, ntrial):
     """this function loads the data, and sets the associated parameters (such as output and input dimensionality and batchsize) according to the specified task, which are either text, music, speech or digits """
     task, data = dictionary['task'], dictionary['data']
 
@@ -689,7 +689,8 @@ def load_data(dictionary):
         L, T, step = 150, 200, 50  
 
         #random.seed( s)
-        Z = sound_set(3)
+        #we pick the set according to trial number 
+        Z = sound_set(3, selector = np.mod(ntrial, 8) + 1 ) 
 
         # Front-end details
         #if hp is None:

### 2017-06-02 main.py
Add files via upload

@@ -91,7 +91,7 @@ def model_driver(d,data):
 def main(dictionary):
 
         ### next thing is determining the hyperparameters
-    np.random.seed( dictionary['seedin'][0] )
+    # np.random.seed( dictionary['seedin'][0] )
     tf.set_random_seed( dictionary['seedin'][1] ) 
     timestamp = str(round(time.time()))
 
@@ -105,20 +105,29 @@ def main(dictionary):
 
     records = [] #information will accumulate in this
     for i in range(dictionary['num_configs']):
-        
+        # np.random.seed(seed = 1234)
         # first get the data and the resulting model parameters  
-        data, parameters = load_data(dictionary = dictionary, ntrial = i)
-        dictionary.update(parameters) # add the necessary model parameters to the dictionary here
-
+        # data, parameters = load_data(dictionary = dictionary)
+        # dictionary.update(parameters) # add the necessary model parameters to the dictionary here
+        # pdb.set_trace()
         while True:  
             try:
-                lr, K, num_layers, momentum = generate_random_hyperparams(
+                # pdb.set_trace()
+                lr, K, num_layers, momentum, set_seed = generate_random_hyperparams(
                         lr_min =  lr_min, lr_max = lr_max, 
                         K_min = K_min, K_max = K_max , 
                         num_layers_min = num_layers_min, 
                         num_layers_max = num_layers_max,
-                        load_hparams = (dictionary['load_hparams'],i))
+                        load_hparams = (dictionary['load_hparams'],i),
+                        num_configs = dictionary['num_configs'])
                 
+                if set_seed:
+                    np.random.seed(1234)
+                
+                data, parameters = load_data(dictionary = dictionary)
+                dictionary.update(parameters) # add the necessary model parameters to the dictionary here
+                
+                    
                 print("Configuration ",i,
                       "K = ",K, 
                       "num_layers = ", num_layers,
@@ -173,13 +182,13 @@ wform = 'full'# either diagonal or full
 input_dictionary = {'seedin' : [1144, 1521], #setting the random seed. First is for numpy, second is for tensorflow 
             'task' : 'source_sep', #this helps us how to load the data with the load_data function in rnns.py 
             'data' : 'timit', #the dataset, options are inside the load_data function 
-            'encoder': 'mb_mod_lstm', #options are: feed_forward, convolutive, mb_mod_lstm
+            'encoder': 'convolutive', #options are: feed_forward, convolutive, mb_mod_lstm
[…]

### 2017-06-02 music_utilities.py

@@ -5,9 +5,9 @@ import pdb
 
 eps = finfo( float32).eps;
 
+# Load a TIMIT data set
 def tset( mf = None, ff = None, dr = None):
     # Where the files are
-    # p = '~/Dropbox/timit/timit-wav/train/' 
     p = '/usr/local/timit/timit-wav/train/'
 
     # Pick a speaker directory
@@ -42,10 +42,10 @@ def tset( mf = None, ff = None, dr = None):
     fs.pop( int(i/10))
     tr = [concatenate(ms), concatenate(fs)]
 
-    return list(map( lambda x : (x-mean(x))/std(x), ts)), list(map( lambda x : (x-mean(x))/std(x), tr))
+    return list(map( lambda x : (x-mean(x))/std(x), ts)), list(map( lambda x : (x-mean(x))/std(x), tr)),mf,ff
 
 
-def sound_set( tp , selector):
+def sound_set( tp):
     import scipy.io.wavfile as wavfile
 
     # Two sinusoids signal
@@ -82,9 +82,9 @@ def sound_set( tp , selector):
 
     # TIMIT male/female set
     elif tp == 3:
-        ts,tr = tset( 'fbjl0', 'mwsh0', selector )
+        # ts,tr = tset( 'fbjl0', 'mwsh0', 5)
         # ts,tr = tset( 'falr0', 'mtqc0', 4)
-        # ts,tr = tset()
+        ts,tr,mf,ff = tset()
         sr = 16000
 
         tr[0] = tr[0][:min(tr[0].shape[0],tr[1].shape[0])]
@@ -109,7 +109,8 @@ def sound_set( tp , selector):
     # Show me
     #soundsc( ts1+ts2, sr)
 
-    return tr1,tr2,ts1,ts2
+    return tr1,tr2,ts1,ts2,mf,ff
+
 
 
 class sound_feats:

### 2017-06-02 rnns.py

@@ -671,16 +671,26 @@ def generate_random_hyperparams(lr_min, lr_max, K_min, K_max, num_layers_min, nu
         Krange = np.tile(Krange, [int(num_configs/len(Krange)),1 ]).transpose()
         Krange = Krange.reshape(-1)
         K = Krange[load_hparams[1]]
+        # pdb.set_trace()
+        set_seed = False
+        if (load_hparams[1]==0):
+            set_seed = True
+        elif (load_hparams[1]>0):
+            # pdb.set_trace()
+            if ( np.abs(Krange[(load_hparams[1])] - Krange[(load_hparams[1]-1)]) > 0):
+                set_seed = True
         
-        lr_exp = np.random.uniform(lr_min, lr_max)
-        lr = 10**(lr_exp)
-        
-        num_layers = np.random.choice(np.arange(num_layers_min, num_layers_max + 1),1)[0]
-        momentum = np.random.uniform(0,1)
+        # lr_exp = np.random.uniform(lr_min, lr_max)
+        # lr = 10**(lr_exp)
+        # num_layers = np.random.choice(np.arange(num_layers_min, num_layers_max + 1),1)[0]
+        # momentum = np.random.uniform(0,1)
+        lr = 0.001
+        num_layers = 1
+        momentum = 0.7
 
-    return lr, K, num_layers, momentum
+    return lr, K, num_layers, momentum, set_seed
 
-def load_data(dictionary, ntrial):
+def load_data(dictionary):
     """this function loads the data, and sets the associated parameters (such as output and input dimensionality and batchsize) according to the specified task, which are either text, music, speech or digits """
     task, data = dictionary['task'], dictionary['data']
 
@@ -690,7 +700,10 @@ def load_data(dictionary, ntrial):
 
         #random.seed( s)
         #we pick the set according to trial number 
-        Z = sound_set(3, selector = np.mod(ntrial, 8) + 1 ) 
+        Z_temp = sound_set(3) 
+        Z = Z_temp[0:4]
+        mf = Z_temp[4]
+        ff = Z_temp[5]
 
         # Front-end details
         #if hp is None:
@@ -755,8 +768,10 @@ def load_data(dictionary, ntrial):
                   'num_buckets':num_buckets,
                   'len_th':len_th,
                   'audio_files':Z,
-                  'FE':FE}
-
+                  'FE':FE,
+                  'mf':mf,
+                  'ff':ff}
+    
     return {'Train1':df_train1, 
             'Train2':df_train2, 
             'Test':df_test, 

### 2017-06-05 paper/paper.tex
Add files via upload

@@
 % ------
-\title{AUTHOR GUIDELINES FOR MLSP PROCEEDINGS MANUSCRIPTS}
+\title{A Neural Network Alternative to Convolutional Audio Models for Source Separation}
 %
@@
 \begin{abstract}
-   In this paper we propose a convolutive and recurrent neural network based extensions to autoencoders for source separation. 
 \end{abstract}
@@
 \begin{keywords}
-One, two, three, four, five
+Auto-encoders, source separation, deep learning.
 \end{keywords}
@@
 \label{sec:intro}
+Recently, several neural network architectures have been proposed to develop algorithms for supervised source separation and speech enhancement.
+(Chandna monaural, Grais Single channel, Park Fully convolutional, Venkataramani end-to-end source).
+Currently, these networks are trained to learn discriminative audio models extensively.
+In other words, the spectrogram of the mixture is given as an input to the network.
+The goal of the network then, is to learn suitable time-frequency masks that separate the input spectrogram into the source and the interference componenents.
+Thus, the networks learn a basis decomposition that often works only for a specific source-interference pair, i.e., these models are not transferable.
+If the interfering signal changes, these networks have to be re-trained to learn suitable models to separate the new interfering signal in the mixture from the source.
 
-Talk about what we are doing and why it is interesting. \\
+A popular technique to learn generative audio models for supervised source separation is the use of Non-negative matrix factorization~(NMF).
+Non-negative matrix factorization~(NMF) approximates a matrix of non-negative elements~$\mathbf{V}\in \mathbb{R^{+}}$ as a product of the basis matrix~$\mathbf{W}$ and the activation matrix~$\mathbf{H}$.
+Here, the matrices $\mathbf{W},~\mathbf{H}$ are non-negative matrices themselves i.e., $\mathbf{W},~\mathbf{H}\in \mathbb{R^{+}}$, where $\mathbb{R^{+}}$ denotes the set of matrices of positive real elements.
+In the case of audio signals, we apply such a factorization on audio spectrograms.
+In this setting, the columns of $\mathbf{W}$ begin to act as representative basis vectors for the source.
+The rows of $\mathbf{H}$ indicate the activity of these basis vectors in time.
+As shown in~~\cite{}(ICASSP Paper reference...), the notion of non-negative audio modeling can be easily generalized by interpreting it as a neural network.
+We can interpret NMF as a non-negative auto-encoder in the following manner,
+\begin{align}
+    \text{$1^{\text{st}}$ layer:~}\mathbf{H} &= g(\mathbf{W^{\ddagger}} \cdot \mathbf{X}) \\
+    \text{$2^{\text{nd}}$ layer:~}\mathbf{X} &= g(\mathbf{W} \cdot \mathbf{H})
+    \label{eq:nmfae}
+\end{align}
 
-Give some background.
+Here, $\mathbf{X}$ represents the input spectrogram, $\mathbf{W^{\ddagger}}$ represents a form of pseudo-inverse of $\mathbf{W}$ and $g(.):\mathbf{R}\rightarrow \mathbf{R^{+}}$ is an element-wise function that maps a real number to the space of positive real numbers.
+As before, the columns of $\mathbf{W}$ act as representative basis vectors and the corresponding rows of $\mathbf{H}$ indicate their respective activations.
+Although non-negativity of the models is not explicitly guaranteed in this formulation, applying a suitable sparsity constraint allows the network to learn suitable non-negative models.
+Additionally, this interpretation enabled a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance.
 
-\section{Autoencoders}
+Spectrograms of speech and audio signals incorporate temporal dependencies that span multiple time frames.
+However, NMF and its neural net equivalent are unable to explicitly utilize these cross-frame patterns available in a spectrogram.
+To alleviate this drawback, Smaragdis~\cite{}(Paris convolutive speech bases) proposed a convolutive version to NMF that allows spectro-temporal patterns as representative basis elements.
+In this paper, we develop a neural network alternative to such convolutive audio models for supervised source separation.
+In doing so, we solve two fundamental problems associated with this task.
+The first step is to develop a suitable neural network architecture to learn convolutive audio models in an adaptive manner.
+Utilizing the models to separate a source from a given mixture forms the second step.
+The remainder of the paper is organized as follows.
+In section~\ref{sec:}, we develop an auto-encoder that can act as an equivalent to conv-NMF audio models.
+Section~\ref{sec:} proposes a novel approach to utilize these models for supervised source separation.
+We evaluate these models in terms of their separation performance in section~\ref{sec:} and conclude in section~\ref{cite:}.
 
-\subsection{Feed-forward Autoencoder} 
-Define standard autoencoder. Provide a toy example to illustrate the shortcomings of this.
-
-\subsection{CNN-CNN Autoencoder} 
+\section{Convolutive NMF as an Auto-encoder}
+\label{sec:conv-nmf}
 The approximation $\hat X$ for a given spectrogram $X$ is computed as follows:
@@
 
-\subsection{Multilayer of CNN-CNN Autoencoder} 
-This loses the interpretability, but maybe good for accuracy?
-
-\subsection{RNN-CNN Autoencoder}
+\section{Extensions to infinite summation}
+\label{sec:rnncnn}
+The convolutive NMF model approximates a non-negative matrix~$mathbf{V}$ as follows,
 
@@
 
+% \section{Multi-layer Extensions}
+% \label{sec:multlayer}
+
+\section{Supervised Source Separation}
+\label{sec:ss}
+Describe the separation procedure.
+
 \section{Experiments}
-
+\label{sec:experiments}
 Try each model in a given $K$ range in speech-speech source separation task.
 
-\section{REFERENCES}
-\label{sec:ref}
+\subsection{Experimental setup}
+\label{subsec:setup}
+
+\subsection{Results and Discussion}
+\label{subsec:results}
+
+\section{Conclusion}
+\label{sec:conclusion}
 
@@
 \bibliographystyle{IEEEbib}
-\bibliography{strings,refs}
+\bibliography{refs.bib}
 

### 2017-06-06 paper/paper.tex
Add files via upload

@@
 \usepackage{amsmath,graphicx,mlspconf}
+\usepackage{amsfonts}
 
@@
 % ------
-\title{A Neural Network Alternative to Convolutional Audio Models for Source Separation}
+\title{A Neural Network Alternative to Convolutive Audio Models for Source Separation}
 %
@@
 If the interfering signal changes, these networks have to be re-trained to learn suitable models to separate the new interfering signal in the mixture from the source.
+\\
 
-A popular technique to learn generative audio models for supervised source separation is the use of Non-negative matrix factorization~(NMF).
-Non-negative matrix factorization~(NMF) approximates a matrix of non-negative elements~$\mathbf{V}\in \mathbb{R^{+}}$ as a product of the basis matrix~$\mathbf{W}$ and the activation matrix~$\mathbf{H}$.
-Here, the matrices $\mathbf{W},~\mathbf{H}$ are non-negative matrices themselves i.e., $\mathbf{W},~\mathbf{H}\in \mathbb{R^{+}}$, where $\mathbb{R^{+}}$ denotes the set of matrices of positive real elements.
+A popular technique to learn generative audio models for supervised source separation is the use of Nnon-negative matrix factorization~(NMF).
+Non-negative matrix factorization~(NMF) matrix of non-negative elements~$\mathbf{X}\in \mathbb{R}_{M\times N}^{\geq0}$ as a product of the basis matrix~$\mathbf{W}$ and the activation matrix~$\mathbf{H}$.
+The notation $\mathbb{R}_{M \times N}^{\geq0}$ represents the set of matrices of non-negative elements of size $M \times N$.
+In this factorization, the basis matrix $\mathbf{W} \in \mathbb{R}_{M \times r}^{\geq{0}}$, the activation matrix $\mathbf{H} \in \mathbb{R}_{r \times N}^{\geq{0}}$ and $r$ represents the rank of the decomposition.
 In the case of audio signals, we apply such a factorization on audio spectrograms.
@@
 The rows of $\mathbf{H}$ indicate the activity of these basis vectors in time.
-As shown in~~\cite{}(ICASSP Paper reference...), the notion of non-negative audio modeling can be easily generalized by interpreting it as a neural network.
+As shown in~~\cite{}(ICASSP Paper reference...), the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network.
 We can interpret NMF as a non-negative auto-encoder in the following manner,
@@
 
-Here, $\mathbf{X}$ represents the input spectrogram, $\mathbf{W^{\ddagger}}$ represents a form of pseudo-inverse of $\mathbf{W}$ and $g(.):\mathbf{R}\rightarrow \mathbf{R^{+}}$ is an element-wise function that maps a real number to the space of positive real numbers.
+Here, $\mathbf{X}$ represents the input spectrogram, $\mathbf{W^{\ddagger}}$ represents a form of pseudo-inverse of $\mathbf{W}$ and $g(.):\mathbf{R}\rightarrow \mathbf{R^{\geq0}}$ is an element-wise function that maps a real number to the space of positive real numbers.
 As before, the columns of $\mathbf{W}$ act as representative basis vectors and the corresponding rows of $\mathbf{H}$ indicate their respective activations.
@@
 Additionally, this interpretation enabled a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance.
+\\
 
@@
 We evaluate these models in terms of their separation performance in section~\ref{sec:} and conclude in section~\ref{cite:}.
+\\
 
-\section{Convolutive NMF as an Auto-encoder}
+\section{Non-negative Convolutiional Auto-encoders}
+\subsection{Network Architecture}
 \label{sec:conv-nmf}
+The convolutive NMF model~\cite{}(Smaragdis convolutive speech bases) approximates a non-negative matrix~$\mathbf{X}\in \mathbb{R}_{M \times N}^{\geq0}$ as,
+\begin{equation}
+    \mathbf{X}(f,t) \approx \sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)
+\end{equation}
+Here, $\mathbf{W}_{i} \in \mathbb{R}_{M \times T}^{\geq0}$ acts as the $i^{\text{th}}$ basis matrix and $\mathbf{H} \in \mathbb{R}_{r \times N}^{\geq0}$ contains the corresponding weights.
+The notation $\mathbf{X}(i,j)$ represents the element of $\mathbf{X}$ indexed by the $i^{\text{th}}$ row and the $j^{\text{th}}$ column.
+Thus, we can interpret this operation as a two-step convolutional auto-encoder as follows,
+
+\begin{align}
+    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= \sum_{i=1}^{r} \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k) \\
+    \text{$2^{\text{nd}}$ layer:~}\hat{\mathbf{X}}(f,t) &= \sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)
+    \label{eq:nmfcnncnn}
+\end{align}
+
+subject to non-negativity of $\mathbf{W}_{i}$ and $\mathbf{H}\geq0$.
+Here, we assume that the convolutional filters~$\mathbf{W},~\mathbf{W}^{\ddagger}$ have a size of $M\times T$ where, $T$ represents the depth of the convolution.
+In this representation, $\mathbf{W}_{i}$ and $\mathbf{H}$ correspond to the $i^{\text{th}}$ basis matrix and the activation matrix respectively.
+The filters of the first convolutional layer act as inverse filters in defining the auto-encoder.
+In the remainder of this section, we will refer to the first convolutional layer as the encoder that estimates a code from the input representation.
+The second convolutional layer generates an approximation of the input from the code and will be referred to as the decoder.
+We can simplify the non-negativity constraints by incorporating a non-linearity into the definitions of the encoder and the decoder.
+Thus,
+
+\begin{align}
+    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= g\left(\sum_{i=1}^{r} \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k)\right) \\
+    \text{$2^{\text{nd}}$ layer:~}\hat{\mathbf{X}}(f,t) &= g\left(\sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)\right)
+    \label{eq:nmfcnncnn}
+\end{align}
+Here, the non-linearity $g(.):\mathbf{R}\rightarrow \mathbf{R^{\geq0}}$ applies an element-wise non-negativity constraint and ensures that the activation matrix and the reconstruction are non-negative.
+We now note a couple of key points about the convolutional auto-encoder.
+
 The approximation $\hat X$ for a given spectrogram $X$ is computed as follows:
@@
 \label{sec:rnncnn}
-The convolutive NMF model approximates a non-negative matrix~$mathbf{V}$ as follows,
 
-This is the same as CNN-CNN case, except the computation of the activations $H$.
+% This is the same as CNN-CNN case, except the computation of the activations $H$.
 In the CNN encoder, each filter was of finite length.
@@
 
-\begin{align}
-	Z(:,k,t) =& \sigma( W Z(:,k,t-1) + U X(:,t)) \notag \\
-	\hat H(k,t) =& \sum_f Z(f,k,t)
-\end{align}
+% \begin{align}
+% 	Z(:,k,t) =& \sigma( W Z(:,k,t-1) + U X(:,t)) \notag \\
+% 	\hat H(k,t) =& \sum_f Z(f,k,t)
+% \end{align}
 
-Give a toy example which shows what this model can do that CNN-CNN can not.
+% Give a toy example which shows what this model can do that CNN-CNN can not.
 

### 2017-06-06 paper/paper.tex
Add files via upload

@@
 \end{align}
-
 subject to non-negativity of $\mathbf{W}_{i}$ and $\mathbf{H}\geq0$.
@@
 \end{align}
-Here, the non-linearity $g(.):\mathbf{R}\rightarrow \mathbf{R^{\geq0}}$ applies an element-wise non-negativity constraint and ensures that the activation matrix and the reconstruction are non-negative.
+Here, the non-linearity $g(.):\mathbb{R}\rightarrow \mathbb{R^{\geq0}}$ applies an element-wise non-negativity constraint and ensures that the activation matrix and the reconstruction are non-negative.
 We now note a couple of key points about the convolutional auto-encoder.
+(i) The output of the encoder gives the latent representation of the decomposition.
+(ii) The weights of the decoder act as the basis vectors of the decomposition.
+(iii) We do not explicitly apply non-negativity constraints on the weights.
+Thus, the basis matrices (decoder filters) can assume negative values.
+
+To train the auto-encoder and learn the basis and activation matrices, in the remainder of the paper, we apply the following approach.
+We minimize the KL-divergence between the input spectrogram~$\mathbf{X}$ and its reconstruction~$\hat{\mathbf{X}}$ given by,
+\begin{equation*}
+    D\left(\mathbf{X},\hat{\mathbf{X}}\right) = \sum_{i,j} \mathbf{X}(i,j)\cdot \text{log}\frac{\mathbf{X}(i,j)}{\hat{\mathbf{X}}(i,j)} - \mathbf{X}(i,j) + \hat{\mathbf{X}}(i,j)
+\end{equation*}
+The network is trained by applying a batch gradient descent training procedure with a batchsize of (...) and the parameters updated using the RMSProp algorithm (**RMSProp reference**), with a learning rate and momentum of (..) and (..) respectively.
 

### 2017-06-06 paper/paper.tex
Add files via upload

@@
 
-A popular technique to learn generative audio models for supervised source separation is the use of Nnon-negative matrix factorization~(NMF).
+A popular technique to learn transferable audio models for supervised source separation is the use of Nnon-negative matrix factorization~(NMF).
 Non-negative matrix factorization~(NMF) matrix of non-negative elements~$\mathbf{X}\in \mathbb{R}_{M\times N}^{\geq0}$ as a product of the basis matrix~$\mathbf{W}$ and the activation matrix~$\mathbf{H}$.
@@
 
-Here, $\mathbf{X}$ represents the input spectrogram, $\mathbf{W^{\ddagger}}$ represents a form of pseudo-inverse of $\mathbf{W}$ and $g(.):\mathbf{R}\rightarrow \mathbf{R^{\geq0}}$ is an element-wise function that maps a real number to the space of positive real numbers.
+Here, $\mathbf{X}$ represents the input spectrogram, $\mathbf{W^{\ddagger}}$ represents a form of pseudo-inverse of $\mathbf{W}$ and $g(.):\mathbf{R}\rightarrow \mathbf{R}^{\geq0}$ is an element-wise function that maps a real number to the space of positive real numbers.
 As before, the columns of $\mathbf{W}$ act as representative basis vectors and the corresponding rows of $\mathbf{H}$ indicate their respective activations.
 Although non-negativity of the models is not explicitly guaranteed in this formulation, applying a suitable sparsity constraint allows the network to learn suitable non-negative models.
-Additionally, this interpretation enabled a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance.
+Additionally, this interpretation enables a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance.
 \\
@@
 Spectrograms of speech and audio signals incorporate temporal dependencies that span multiple time frames.
-However, NMF and its neural net equivalent are unable to explicitly utilize these cross-frame patterns available in a spectrogram.
+However, NMF and its neural network equivalent are unable to explicitly utilize these cross-frame patterns available in a spectrogram.
 To alleviate this drawback, Smaragdis~\cite{}(Paris convolutive speech bases) proposed a convolutive version to NMF that allows spectro-temporal patterns as representative basis elements.
@@
 
-The approximation $\hat X$ for a given spectrogram $X$ is computed as follows:
+\subsection{Practical Considerations}
+\label{subsec:practical}
+Having developed the convolutional auto-encoder equivalent to convolutive NMF, we can now begin to understand the nature of the basis and activation matrices learned by the network.
+To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{}(**Figure reference**)
 
-\begin{align}
-	\hat H(k,t) =& \sigma_1\left ( \sum_{f,t'} X(f,t-t') F_e(f,t',k) \right ) \notag \\
-	\hat X(f,t) =& \sigma_2 \left ( \sum_k \sum_{t'} \hat H(k,t-t') F_d(f, t',k) \right )
-\end{align}
+% The approximation $\hat X$ for a given spectrogram $X$ is computed as follows:
+
+% \begin{align}
+% 	\hat H(k,t) =& \sigma_1\left ( \sum_{f,t'} X(f,t-t') F_e(f,t',k) \right ) \notag \\
+% 	\hat X(f,t) =& \sigma_2 \left ( \sum_k \sum_{t'} \hat H(k,t-t') F_d(f, t',k) \right )
+% \end{align}
 

### 2017-06-07 paper/paper.tex
Add files via upload

@@
 \end{align}
-subject to non-negativity of $\mathbf{W}_{i}$ and $\mathbf{H}\geq0$.
+subject to non-negativity of $\mathbf{W}_{i}$ and $\mathbf{H}$.
 Here, we assume that the convolutional filters~$\mathbf{W},~\mathbf{W}^{\ddagger}$ have a size of $M\times T$ where, $T$ represents the depth of the convolution.
@@
 \end{align}
-Here, the non-linearity $g(.):\mathbb{R}\rightarrow \mathbb{R^{\geq0}}$ applies an element-wise non-negativity constraint and ensures that the activation matrix and the reconstruction are non-negative.
+Here, the non-linearity $g(.):\mathbb{R}\rightarrow \mathbb{R}^{\geq0}$ applies an element-wise non-negativity constraint and ensures that the activation matrix and the reconstruction are non-negative.
 We now note a couple of key points about the convolutional auto-encoder.
@@
 Having developed the convolutional auto-encoder equivalent to convolutive NMF, we can now begin to understand the nature of the basis and activation matrices learned by the network.
-To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{}(**Figure reference**)
+To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{}(**Figure reference**)7
 
@@
 \label{sec:ss}
-Describe the separation procedure.
+The problem of supervised source separation is solved as a two-step procedure~\cite{}(**Paris supervised and semi-supervised**).
+The fist step of the procedure is to learn suitable models for a given source.
+We refer to this step as the training step.
+In the second step, we use these models to explain the contribution of the source in an unknown mixture.
+In section~\ref{sec:conv-nmf}, we have developed the auto-encoder architecture to learn suitable convolutive models for a given source.
+We now turn our attention to the problem of using the models for separating the source in an unknown mixture.
+
+The previous approach to source separation~\cite{}(**ICASSP paper**) involves estimating the latent representation of the bases, which captures the contribution of the bases to each frame of the mixture spectrogram.
+This does not utilize the encoder component of the auto-encoder for the separation task.
+In this paper, we present a novel-separation scheme that utilizes the complete auto-encoder architecture for separation.
+We do so by using the following setup for separation.
+Given an input spectrogram $\mathbf{X}$, the auto-encoder produces an approximation that models the spectrogram in terms of its weights.
+We will denote to this approximation as,
+\begin{equation}
+    \hat{\mathbf{X}} = Ae(\mathbf{X}|\theta)
+    \label{eq:separation_ae}
+\end{equation}
+Here, $\theta$ denotes the weights (parameters) of the auto-encoder.
+Thus, given the trained auto-encoders for sources, i.e., given $\theta_{1}$ and $\theta_{2}$, the goal of separation is to identify suitable spectrograms $\mathbf{X}_{1}$ and $\mathbf{X}_{2}$ such that,
+\begin{equation}
+    \mathbf{X}_{m} = Ae(\mathbf{X}_{1}|\theta_{1}) + Ae(\mathbf{X}_{2}|\theta_{2}) 
+    \label{eq:separation}
+\end{equation}
+Here, $\mathbf{X}_{m}$ represents the spectrogram of the mixture.
+In other words, the separation procedure attempts to estimate the mixture spectrogram as the sum of spectrograms of the individual sources.
+This assumption of additivity of spectrograms is similar to the separation procedure used in~\cite{}(**NMF, Icassp**).
+However, we directly estimate the source spectrograms without estimating the latent representation.
+We train the network defined by~(\ref{eq:separation}) for an appropriate input instead of training for the weights of the network.
+As before, we minimize the KL divergence between mixture spectrogram $\mathbf{X}_{m}$ and its approximation~$\mathbf{X}_{1} + \mathbf{X}_{2}$.
+Conceptually, this problem is not different to training a neural network.
+The equivalence can be seen by applying a transposition to the auto-encoder definitions in (\ref{eq:nmfcnncnn}).
+This also allows a generalized separation procedure to be applied even when the underlying architectures of the auto-encoders are changed.
 

### 2017-06-09 main.py
I have added the toy_example option (see load_data function for custom toy example)

@@ -24,73 +24,93 @@ def model_driver(d,data):
     config.gpu_options.allow_growth=True
     config.allow_soft_placement=True
 
+    if d['task'] == 'source_sep':
+        
+        #build graph 1
+        with tf.variable_scope("model1"):
+            rnn1 = rnn( model_specs = d, initializer = d['init'], model_num = 1)
+            rnn1_handles = rnn1.build_graph()  
+        
+        with tf.Session(config = config) as sess:
+            sess.run(tf.initialize_all_variables())
+            all_times, tr_logls, _, _ = rnn1.optimizer(data = data['Train1'], 
+                    rnn_handles = rnn1_handles, sess = sess)
+            
+            vars_np1 = rnn1.save_modelvars_np(sess)
+
+        tf.reset_default_graph()
+
+        #build graph 2
+        with tf.variable_scope("model2"):
+            rnn2 = rnn( model_specs = d, initializer = d['init'], model_num = 2)
+            rnn2_handles = rnn2.build_graph()  
+
+        with tf.Session(config = config) as sess:
+            sess.run(tf.initialize_all_variables())
+            all_times, tr_logls, _, _ = rnn2.optimizer(data = data['Train2'], 
+                    rnn_handles = rnn2_handles, sess = sess) 
+
+            vars_np2 = rnn2.save_modelvars_np(sess)
+
+        tf.reset_default_graph()
+
+        all_vars_np = vars_np1 + vars_np2
+
+        rnn_sep_opt = rnn( model_specs = d, initializer = all_vars_np, input_opt = True)
+        rnn_sep_opt_handles = rnn_sep_opt.build_separation_graph()
+
+        with tf.Session(config = config) as sess:
+            sess.run(tf.initialize_all_variables())
+        
+            x1hat, x2hat = rnn_sep_opt.input_optimizer(data = data['Test'], 
+                                        rnn_handles = rnn_sep_opt_handles, sess = sess) 
+
+        Z = d['audio_files']
+        mixture = np.concatenate(data['Test']['data'].values, axis = 1)    
+        phase = np.concatenate(data['Test']['phase'].values, axis = 1)    
+        
+        x1hat = x1hat[:,:mixture.shape[1]]
+        x2hat = x2hat[:,:mixture.shape[1]]
+        eps = 1e-20
+
+        o1 = d['FE'].ife( x1hat * (mixture/(x1hat+x2hat+eps)), phase)
[…]

### 2017-06-09 rnns.py

@@ -248,10 +248,9 @@ class rnn(object):
             elif d['optimizer'] == 'Adadelta':
                 train_step = tf.train.AdadeltaOptimizer(d['LR']).minimize(cost)   
 
-            if d['task'] == 'source_sep':
-                relevant_inds = tf.squeeze(tf.where(tf.cast(mask,tf.bool)))
-                preds = tf.gather(yhat,relevant_inds) 
-                targets = tf.gather(y,relevant_inds) 
+            relevant_inds = tf.squeeze(tf.where(tf.cast(mask,tf.bool)))
+            preds = tf.gather(yhat,relevant_inds) 
+            targets = tf.gather(y,relevant_inds) 
             
             #return the graph handles 
             graph_handles = {'train_step':train_step,
@@ -406,7 +405,6 @@ class rnn(object):
             yhat = (yhat + b)
             
             return yhat
-
 
     def build_separation_graph(self):
         d = self.model_specs #unpack the model specifications
@@ -695,7 +693,36 @@ def load_data(dictionary):
     task, data = dictionary['task'], dictionary['data']
 
     #data = [generate_separationdata(L,T,step)]
-    if task == 'source_sep':
+    if task == 'toy_example':
+        
+        # load the toy data as M
+        # put the length of M in lengths 1
+        L, T, step = 10, 100, 25
+        
+        M = [generate_separationdata(L,T,step)]
+        lengths = [T]
+       
+        # set the toy magnitude spectrogram as M
+        d = {'data':M, 'lengths': lengths}
+        df_train = pd.DataFrame( d )
+        
+        L1 = L2 = L
+        outstage = 'relu'
+        mapping_mode = 'seq2seq'
+        iterator = 'SimpleDataIterator'
+
+        parameters = {'batchsize1':T,
+                      'L1':L1,'L2':L2,
+                      'outstage':outstage,
+                      'mapping_mode':mapping_mode,
+                      'iterator':iterator,
+                      'num_buckets':None
+                      }
+        
+        return df_train, parameters
+
+
+    elif task == 'source_sep':
         L, T, step = 150, 200, 50  
 
         #random.seed( s)
@@ -738,44 +765,49 @@ def load_data(dictionary):
 
         #mixtures
         M_t, P_t = FE.fe( Z[2]+Z[3] )
-        M_t, P_t = [M_t], [P_t]
-        #M_t = np.split(M_t, np.arange(len_th, M_t.shape[1],len_th), axis = 1)
[…]

### 2017-06-09 paper/paper.tex

@@
 \label{sec:intro}
-Recently, several neural network architectures have been proposed to develop algorithms for supervised source separation and speech enhancement.
+Several neural network architectures have been proposed to develop algorithms for supervised source separation and speech enhancement.
 (Chandna monaural, Grais Single channel, Park Fully convolutional, Venkataramani end-to-end source).
@@
 Having developed the convolutional auto-encoder equivalent to convolutive NMF, we can now begin to understand the nature of the basis and activation matrices learned by the network.
-To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{}(**Figure reference**)7
+To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{}(**Figure reference**).
 
@@
 Here, $\theta$ denotes the weights (parameters) of the auto-encoder.
-Thus, given the trained auto-encoders for sources, i.e., given $\theta_{1}$ and $\theta_{2}$, the goal of separation is to identify suitable spectrograms $\mathbf{X}_{1}$ and $\mathbf{X}_{2}$ such that,
+Thus, given the trained auto-encoders, i.e., given $\theta_{1}$ and $\theta_{2}$, the goal of separation is to identify suitable spectrograms $\mathbf{X}_{1}$ and $\mathbf{X}_{2}$ such that,
 \begin{equation}
@@
 \end{equation}
-Here, $\mathbf{X}_{m}$ represents the spectrogram of the mixture.
+Here, $\mathbf{X}_{m}$ represents the spectrogram of the mixture and $\mathbf{X}_{1}$,~$\mathbf{X}_{2}$ denote the separated source spectrograms.
 In other words, the separation procedure attempts to estimate the mixture spectrogram as the sum of spectrograms of the individual sources.
@@
 
+Having obtained the contributions of the sources (separated spectrograms), the next step is to transform these spectrograms back into the time domain. This is given as,
+\begin{equation}
+    x_{i}(t) = \text{STFT}^{-1}\left(\frac{\mathbf{X_{i}}}{\sum_{i}\mathbf{X}_{i}}\odot \mathbf{X}_{m} \odot e^{i\Phi_{m}}\right)~\text{for}~i\in\{1, 2\}
+\end{equation}
+Here $x_{i}(t)$ denotes the separated speech signal in time and $\Phi_{m}$ represents the phase of the mixture.
+Also, $\odot$ represents the element-wise multiplication operation and the division is also element-wise.
+
 \section{Experiments}
 \label{sec:experiments}
-Try each model in a given $K$ range in speech-speech source separation task.
+We now describe the experiments used to evaluate our auto-encoder based convolutive audio models.
+We do so by comparing the separation performance of these models to the neural network based models proposed in~\cite{}(**ICASSP paper**).
+For a uniform experimental setup, we apply a separation scheme described in~\ref{sec:ss} to both the models.
+We compare the performance in terms of median BSS\_eval metrics~\cite{}(**BSS eval paper**) viz., signal-to-distortion (SDR), signal-to-interference (SIR) and signal-to-artifact ratio (SAR) parameters.
+We also compare the median intelligibility scores in terms of the STOI index~\cite{}(**STOI paper**).
+To compare these models, we use the TIMIT corpus~\cite{} (**Timit reference**).
+To form these mixtures, we randomly select a pair of male-female speakers from the TIMIT corpus.
+Of the $10$ utterances available for each speaker, one utterance is randomly selected for each speaker.
+These two selected utterances are mixed at $0 dB$ to generate the testing mixture.
+For the evaluation, we generate $20$ such mixtures and compare the models for different parameter configurations.
+As the pre-processing step, we apply a $1024$ point short-time Fourier transform representation with a hop of $25\%$.
+The magnitude spectrogram is then given as an input to the network.
+
+We perform these evaluations at multiple parameter configurations with varying values for decomposition rank and filter size.
+
+*** Talk about how you vary or select the values of K and T***
 

### 2017-06-10 paper/paper.tex
Add files via upload

@@
 
+\begin{figure}[t]
+\centering
+  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\linewidth]{Figs/CNN_demo.png}
+  \caption{...}~\label{fig:cnn_demo}
+\end{figure}
+
 \subsection{Practical Considerations}
@@
 \end{equation}
-Here $x_{i}(t)$ denotes the separated speech signal in time and $\Phi_{m}$ represents the phase of the mixture.
+Here $x_{i}(t)$ denotes the separated speech signal in time and $\Phi_{m}$ represents the phase of the mixture and $\text{STFT}^{-1}$ is the inverse short-time Fourier transform operation that transforms the complex spectrogram into its corresponding time domain representation.
 Also, $\odot$ represents the element-wise multiplication operation and the division is also element-wise.
+
+\begin{figure*}[ht!]
+\centering
+  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\linewidth]{Figs/SeparationPerformance.png}
+  \caption{...}~\label{fig:separation_performance}
+\end{figure*}
 
@@
 \label{sec:experiments}
-We now describe the experiments used to evaluate our auto-encoder based convolutive audio models.
+We now describe the experimental setup used to evaluate our auto-encoder based convolutive audio models.
 We do so by comparing the separation performance of these models to the neural network based models proposed in~\cite{}(**ICASSP paper**).
@@
 
-\subsection{Experimental setup}
-\label{subsec:setup}
-
 \subsection{Results and Discussion}

### 2017-06-10 paper/paper.tex
I made some suggestions for the paper

@@
 In other words, the spectrogram of the mixture is given as an input to the network.
-The goal of the network then, is to learn suitable time-frequency masks that separate the input spectrogram into the source and the interference componenents.
+The goal of the network then, is to learn suitable time-frequency masks that separate the input spectrogram into the source and the interference components.
 Thus, the networks learn a basis decomposition that often works only for a specific source-interference pair, i.e., these models are not transferable.
 If the interfering signal changes, these networks have to be re-trained to learn suitable models to separate the new interfering signal in the mixture from the source.
-\\
+\textbf{Cem: I am not sure if starting the paper with a generative vs discriminative (transferability) discussion is the right way.
+}  \\
 
-A popular technique to learn transferable audio models for supervised source separation is the use of Nnon-negative matrix factorization~(NMF).
+A popular technique to learn transferable audio models for supervised source separation is the use of Non-negative matrix factorization~(NMF).
 Non-negative matrix factorization~(NMF) matrix of non-negative elements~$\mathbf{X}\in \mathbb{R}_{M\times N}^{\geq0}$ as a product of the basis matrix~$\mathbf{W}$ and the activation matrix~$\mathbf{H}$.
@@
 In this factorization, the basis matrix $\mathbf{W} \in \mathbb{R}_{M \times r}^{\geq{0}}$, the activation matrix $\mathbf{H} \in \mathbb{R}_{r \times N}^{\geq{0}}$ and $r$ represents the rank of the decomposition.
-In the case of audio signals, we apply such a factorization on audio spectrograms.
-In this setting, the columns of $\mathbf{W}$ begin to act as representative basis vectors for the source.
+In the case of audio signals, we apply NMF on audio spectrograms.
+In the audio setting, the columns of $\mathbf{W}$ begin to act as representative basis vectors for the source.
 The rows of $\mathbf{H}$ indicate the activity of these basis vectors in time.
-As shown in~~\cite{}(ICASSP Paper reference...), the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network.
+\textbf{Cem: I am not sure how smooth the transition to neural nets here.} As shown in~~\cite{}(ICASSP Paper reference...), the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network.
 We can interpret NMF as a non-negative auto-encoder in the following manner,
@@
 As before, the columns of $\mathbf{W}$ act as representative basis vectors and the corresponding rows of $\mathbf{H}$ indicate their respective activations.
-Although non-negativity of the models is not explicitly guaranteed in this formulation, applying a suitable sparsity constraint allows the network to learn suitable non-negative models.
+Although non-negativity of the models \textbf{Do you mean non-negativity of the parameters?} is not explicitly guaranteed in this formulation, applying a suitable sparsity constraint allows the network to learn suitable non-negative models.
 Additionally, this interpretation enables a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance.

### 2017-06-11 main.py
I have fixed the rnn encoder

@@ -96,7 +96,6 @@ def model_driver(d,data):
             rnn1 = rnn( model_specs = d, initializer = d['init'], model_num = 1)
             rnn1_handles = rnn1.build_graph()  
         
-        pdb.set_trace()
         with tf.Session(config = config) as sess:
             sess.run(tf.initialize_all_variables())
             all_times, tr_logls, _, _ = rnn1.optimizer(data = data, 
@@ -160,7 +159,7 @@ def main(dictionary):
                 try: # Sometimes resources may get exhausted, this exception handles that 
                     dictionary.update({'LR': lr, 
                                         'K': K,
-                                        'K_in':10, # include this in search later
+                                        'K_in': 9, # include this in search later
                                         'num_layers': num_layers,
                                         'min_params': min_params,
                                         'max_params': max_params,
@@ -198,11 +197,11 @@ def main(dictionary):
 
 
 #import matplotlib.pyplot as plt
-wform = 'full'# either diagonal or full
+wform = 'block_diagonal'# either diagonal, block_diagonal or full
 input_dictionary = {'seedin' : [1144, 1521], #setting the random seed. First is for numpy, second is for tensorflow 
-            'task' : 'toy_example', #this helps us how to load the data with the load_data function in rnns.py 
+            'task' : 'source_sep', #this helps us how to load the data with the load_data function in rnns.py 
             'data' : 'timit', #the dataset, options are inside the load_data function 
-            'encoder': 'convolutive', #options are: feed_forward, convolutive, mb_mod_lstm
+            'encoder': 'mb_mod_lstm', #options are: feed_forward, convolutive, mb_mod_lstm
             'decoder': 'convolutive',
             'wform' : wform, 
             'wform_global' : wform,
@@ -210,7 +209,7 @@ input_dictionary = {'seedin' : [1144, 1521], #setting the random seed. First is 
             'start' : 0,  #this is used to start from a certain point (can be useful with fixed seed, or when hyper-parameters are loaded) 
             'EP' : 2000, #number of epochs per run 
             'dropout' : [1, 1], #first is the input second is the output keep probability 
-            'device' : 'gpu:2', #the device to be used in the computations 
+            'device' : 'gpu:0', #the device to be used in the computations 
             'server': socket.gethostname(),
             'verbose': True, #this prints out the batch location
             'load_hparams': True, #this loads hyper-parameters from a results file

### 2017-06-11 rnns.py

@@ -17,8 +17,9 @@ from music_utilities import *
 class ModLSTMCell(tf.contrib.rnn.RNNCell):
     """Modified LSTM Cell """
 
-    def __init__(self, num_units, initializer = tf.contrib.layers.xavier_initializer(uniform=True, seed=2, dtype=tf.float32), wform = 'diagonal', input_opt = False, model_num = None):
+    def __init__(self, num_units, initializer = tf.contrib.layers.xavier_initializer(uniform=True, seed=2, dtype=tf.float32), wform = 'diagonal', input_opt = False, model_num = None, num_rnns = 1):
         self._num_units = num_units
+        self._num_rnns = num_rnns
         self.init = initializer
         self.wform = wform 
         self.input_opt = input_opt
@@ -26,11 +27,12 @@ class ModLSTMCell(tf.contrib.rnn.RNNCell):
 
     @property
     def state_size(self):
-        return tf.contrib.rnn.LSTMStateTuple(self._num_units, self._num_units)
+        return tf.contrib.rnn.LSTMStateTuple(self._num_units*self._num_rnns, 
+                                             self._num_units*self._num_rnns)
 
     @property
     def output_size(self):
-        return self._num_units
+        return self._num_units*self._num_rnns
 
     def __call__(self, inputs, state, scope=None):
         with tf.variable_scope(scope or type(self).__name__):  # "GRUCell"
@@ -38,14 +40,34 @@ class ModLSTMCell(tf.contrib.rnn.RNNCell):
             c, h = state
             init = self.init
             self.L1 = inputs.get_shape().as_list()[1]
-           
-            mats, biases = self.get_params_parallel()
-            if self.wform == 'full' or self.wform == 'diag_to_full':
-                
+            
+            if self.wform == 'full':
+                mats, biases = self.get_params_parallel()
+            elif self.wform == 'block_diagonal':
+                Ws, Us, biases = self.get_params_parallel()
+            
+            if self.wform == 'full':                 
                 res = tf.matmul(tf.concat([h,inputs],axis=1),mats)
                 res_wbiases = tf.nn.bias_add(res, biases)
            
-                i,j,f,o = tf.split(res_wbiases,num_or_size_splits=4,axis=1) 
+                i,j,f,o = tf.split(res_wbiases, num_or_size_splits=4, axis=1) 
+            elif self.wform == 'block_diagonal':
+                h = tf.split(h, num_or_size_splits = int(self._num_rnns), 
+                            axis=1)  
[…]

### 2017-06-11 paper/paper.tex

@@
 \begin{abstract}
+We present a convolutiional auto-encoder that can act as a neural network equivalent to convolutive audio models.
+We demonstrate the ability of the network to learn optimal bases that resemble phoneme magnitude-spectra and also show how it can be used for supervised source separation of speech mixtures.
+In terms of separation performance, the ability of the network to learn cross-frame characteristics from input spectrograms enables these models to significantly outperform their feed-forward counterpart.
+More importantly, we can now exploit the available diversity of neural networks to propose novel and potentially powerful extensions to existing auto-encoder architectures, based on a cascade of recurrent and convolutional neural network layers.
+
 \end{abstract}
@@
 \begin{keywords}
-Auto-encoders, source separation, deep learning.
+Auto-encoders, source separation, deep learning, convolutive models.
 \end{keywords}
@@
 \label{sec:intro}
-Several neural network architectures have been proposed to develop algorithms for supervised source separation and speech enhancement.
-(Chandna monaural, Grais Single channel, Park Fully convolutional, Venkataramani end-to-end source).
+Several neural network architectures have been proposed to develop algorithms for supervised source separation and speech enhancement~\cite{chandna2017monoaural,grais2017single, venkataramani2017end, park2016fully}.
 Currently, these networks are trained to learn discriminative audio models extensively.
@@
 The rows of $\mathbf{H}$ indicate the activity of these basis vectors in time.
-\textbf{Cem: I am not sure how smooth the transition to neural nets here.} As shown in~~\cite{}(ICASSP Paper reference...), the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network.
+\textbf{Cem: I am not sure how smooth the transition to neural nets here.} As shown in~~\cite{smaragdis2017aneural}, the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network.
 We can interpret NMF as a non-negative auto-encoder in the following manner,
@@
 As before, the columns of $\mathbf{W}$ act as representative basis vectors and the corresponding rows of $\mathbf{H}$ indicate their respective activations.
-Although non-negativity of the models \textbf{Do you mean non-negativity of the parameters?} is not explicitly guaranteed in this formulation, applying a suitable sparsity constraint allows the network to learn suitable non-negative models.
+Although non-negativity of the network-parameters (models) is not explicitly guaranteed in this formulation, applying a suitable sparsity constraint allows the network to learn suitable non-negative models.
 Additionally, this interpretation enables a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance.
@@
 However, NMF and its neural network equivalent are unable to explicitly utilize these cross-frame patterns available in a spectrogram.
-To alleviate this drawback, Smaragdis~\cite{}(Paris convolutive speech bases) proposed a convolutive version to NMF that allows spectro-temporal patterns as representative basis elements.
+To alleviate this drawback, Smaragdis~\cite{smaragdis2007convolutive} proposed a convolutive version to NMF that allows spectro-temporal patterns as representative basis elements.
 In this paper, we develop a neural network alternative to such convolutive audio models for supervised source separation.
@@
 \label{sec:conv-nmf}
-The convolutive NMF model~\cite{}(Smaragdis convolutive speech bases) approximates a non-negative matrix~$\mathbf{X}\in \mathbb{R}_{M \times N}^{\geq0}$ as,
+The convolutive NMF model~\cite{smaragdis2007convolutive} approximates a non-negative matrix~$\mathbf{X}\in \mathbb{R}_{M \times N}^{\geq0}$ as,
 \begin{equation}
@@
 The notation $\mathbf{X}(i,j)$ represents the element of $\mathbf{X}$ indexed by the $i^{\text{th}}$ row and the $j^{\text{th}}$ column.
-Thus, we can interpret this operation as a two-step convolutional auto-encoder as follows,
+Thus, we can interpret this operation as a two-step convolutional auto-encoder~(CAE) as follows,
 
@@
 In this representation, $\mathbf{W}_{i}$ and $\mathbf{H}$ correspond to the $i^{\text{th}}$ basis matrix and the activation matrix respectively.
-The filters of the first convolutional layer act as inverse filters in defining the auto-encoder.
+The filters of the first convolutional neural network~(CNN) act as inverse filters in defining the auto-encoder.
 In the remainder of this section, we will refer to the first convolutional layer as the encoder that estimates a code from the input representation.
-The second convolutional layer generates an approximation of the input from the code and will be referred to as the decoder.
+The second CNN layer generates an approximation of the input from the code and will be referred to as the decoder.
 We can simplify the non-negativity constraints by incorporating a non-linearity into the definitions of the encoder and the decoder.
@@
 \end{align}
-Here, the non-linearity $g(.):\mathbb{R}\rightarrow \mathbb{R}^{\geq0}$ applies an element-wise non-negativity constraint and ensures that the activation matrix and the reconstruction are non-negative.
-We now note a couple of key points about the convolutional auto-encoder.
+Here, the non-linearity $g(.):\mathbb{R}\rightarrow \mathbb{R}^{\geq0}$ applies an element-wise non-linearity and ensures that the activation matrix and the reconstruction are non-negative.
+We now note a couple of key points about the CAE.
 (i) The output of the encoder gives the latent representation of the decomposition.
@@
 \end{equation*}
-The network is trained by applying a batch gradient descent training procedure with a batchsize of (...) and the parameters updated using the RMSProp algorithm (**RMSProp reference**), with a learning rate and momentum of (..) and (..) respectively.
 
@@
 \label{subsec:practical}
-Having developed the convolutional auto-encoder equivalent to convolutive NMF, we can now begin to understand the nature of the basis and activation matrices learned by the network.
-To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{}(**Figure reference**).
+Having developed the CAE equivalent to convolutive NMF, we can now begin to understand the nature of the basis and activation matrices learned by the network.
+To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{cnn_demo}.
+We also incorporate sparsity constraints on the activation, i.e., the output of the first CNN.
+We see that the CNN filter resembles a snippet of the input spectrogram.
+From the nature of the activation, it is clear that the encoder acts as a matched-filter and identifies the points in time when the corresponding pattern becomes active.
+As shown in (\ref{eq:nmfcnncnn}), the time-frequency pattern is captured by the filters of the decoder.
+Thus, the encoder learns the inverse-filter of the decoder.
 
@@
 \label{sec:ss}
-The problem of supervised source separation is solved as a two-step procedure~\cite{}(**Paris supervised and semi-supervised**).
+The problem of supervised source separation is solved as a two-step procedure~\cite{smaragdis2007supervised}.
 The fist step of the procedure is to learn suitable models for a given source.
@@
 
-The previous approach to source separation~\cite{}(**ICASSP paper**) involves estimating the latent representation of the bases, which captures the contribution of the bases to each frame of the mixture spectrogram.
+The previous approach to source separation~\cite{smaragdis2017aneural} using neural-network based audio-models involves estimating the latent representation of the bases, which captures the contribution of the bases to each frame of the mixture spectrogram.
 This does not utilize the encoder component of the auto-encoder for the separation task.
@@
 We do so by using the following setup for separation.
-Given an input spectrogram $\mathbf{X}$, the auto-encoder produces an approximation that models the spectrogram in terms of its weights.
-We will denote to this approximation as,
+
+Given an input spectrogram $\mathbf{X}$, the auto-encoder produces an approximation of the input spectrogram which is a linear combination of its weights. We will denote to this approximation as,
 \begin{equation}
@@
 Here, $\theta$ denotes the weights (parameters) of the auto-encoder.
-Thus, given the trained auto-encoders, i.e., given $\theta_{1}$ and $\theta_{2}$, the goal of separation is to identify suitable spectrograms $\mathbf{X}_{1}$ and $\mathbf{X}_{2}$ such that,
+For the separation procedure, given the trained auto-encoders, (i.e., given $\theta_{1}$ and $\theta_{2}$), the goal is to identify suitable input spectrograms $\mathbf{X}_{1}$ and $\mathbf{X}_{2}$ such that,
 \begin{equation}
@@
 \end{equation}
-Here, $\mathbf{X}_{m}$ represents the spectrogram of the mixture and $\mathbf{X}_{1}$,~$\mathbf{X}_{2}$ denote the separated source spectrograms.
-In other words, the separation procedure attempts to estimate the mixture spectrogram as the sum of spectrograms of the individual sources.
-This assumption of additivity of spectrograms is similar to the separation procedure used in~\cite{}(**NMF, Icassp**).
-However, we directly estimate the source spectrograms without estimating the latent representation.
-We train the network defined by~(\ref{eq:separation}) for an appropriate input instead of training for the weights of the network.
+In this equation, $\mathbf{X}_{m}$ represents the spectrogram of the mixture and $\mathbf{X}_{1}$,~$\mathbf{X}_{2}$ denote the separated source spectrograms.
+Thus, similar to NMF, this approach assumes that the magnitude-spectrogram of the mixture is the sum of magnitude-spectrograms of the underlying sources.
+However, in this separation procedure, we directly estimate the source magnitude-spectrograms without estimating the latent representation.
+To do so, we train the network defined by~(\ref{eq:separation}) for an appropriate input $\mathbf{X}_{1}$,~$\mathbf{X}_{2}$, instead of training for the weights of the network.
 As before, we minimize the KL divergence between mixture spectrogram $\mathbf{X}_{m}$ and its approximation~$\mathbf{X}_{1} + \mathbf{X}_{2}$.
@@
 We now describe the experimental setup used to evaluate our auto-encoder based convolutive audio models.
-We do so by comparing the separation performance of these models to the neural network based models proposed in~\cite{}(**ICASSP paper**).
-For a uniform experimental setup, we apply a separation scheme described in~\ref{sec:ss} to both the models.
-We compare the performance in terms of median BSS\_eval metrics~\cite{}(**BSS eval paper**) viz., signal-to-distortion (SDR), signal-to-interference (SIR) and signal-to-artifact ratio (SAR) parameters.
-We also compare the median intelligibility scores in terms of the STOI index~\cite{}(**STOI paper**).
-To compare these models, we use the TIMIT corpus~\cite{} (**Timit reference**).
-To form these mixtures, we randomly select a pair of male-female speakers from the TIMIT corpus.
+We construct a set of training and test examples using the TIMIT corpus~\cite{timit} for the evaluation.
+To form the examples, we randomly select a pair of male-female speakers from the TIMIT corpus.
 Of the $10$ utterances available for each speaker, one utterance is randomly selected for each speaker.
-These two selected utterances are mixed at $0 dB$ to generate the testing mixture.
+These two selected utterances are mixed at $0~dB$ to generate the testing mixture.
+The remaining $9$ utterances are used as training data to construct models for the sources.
+In other words, these examples are used to train the respective (convolutional/ feed-forward) auto-encoders.
 For the evaluation, we generate $20$ such mixtures and compare the models for different parameter configurations.
-As the pre-processing step, we apply a $1024$ point short-time Fourier transform representation with a hop of $25\%$.
+As a pre-processing step, we apply a $1024$ point short-time Fourier transform representation with a hop of $25\%$.
 The magnitude spectrogram is then given as an input to the network.
 
-We perform these evaluations at multiple parameter configurations with varying values for decomposition rank and filter size.
+The neural networks are initialized using the Xavier initialization scheme~\cite{glorot2010understanding}.
+The networks are trained by applying a batch gradient descent training procedure with a batchsize of $18$ frames and the parameters updated using the RMSProp algorithm~\cite{tieleman2012rmsprop}, with a learning rate and momentum of $0.001$ and $0.7$ respectively.
 
-*** Talk about how you vary or select the values of K and T***
+The CNN filters are selected to be $512$-point tall and $8$-point wide.
+Thus, the convolutions are performed only along the time axis.
+The number of CNN filters also decides the number of components in the decomposition.
+We evaluate these models over a varying number of CNN filters ranging from $10$ to $100$ uniformly in steps of $10$.
+We compare the separation performance in terms of median BSS\_eval metrics~\cite{fevotte2005bss_eval} viz., signal-to-distortion (SDR), signal-to-interference (SIR) and signal-to-artifact ratio (SAR) parameters.
 
@@
 \label{subsec:results}
+Figure~\ref{fig:separation_performance} gives the separation performance for the CNN-CNN models (left) and RNN-CNN models (right) for varying values of number of filters~$K$.
+We evaluate the proposed CAE models by comparing the separation performance to their equivalent feed-forward~(FF) counterparts proposed in~\cite{smaragdis2017aneural}.
+In order to maintain a uniform experimental setup, we apply the separation scheme described in~\ref{sec:ss} to all the models.
+We plot the results in terms of a violin plot that also indicates the median value and the corresponding inter-quartile range for each $K$.
+
+We see that the CAE models significantly out-perform their corresponding FF versions.
+This can be seen from the fact that the inter-quartile range in SDR for CAE models is higher than the inter-quartile range of corresponding FF models for several values of $K$.
+This improvement is a consequence of reduced interference generated by these models in source separation (See SIR plots).
+Although the SAR for CAE models degrades slightly as compared to FF models, the improvement in significant improvement in SIR compensates for this loss.
 

### 2017-06-11 paper/paper.tex
Introduction restructured, CNN separation performance figure.

@@
 We present a convolutiional auto-encoder that can act as a neural network equivalent to convolutive audio models.
-We demonstrate the ability of the network to learn optimal bases that resemble phoneme magnitude-spectra and also show how it can be used for supervised source separation of speech mixtures.
+We demonstrate the ability of the network to learn optimal bases that resemble phoneme magnitude-spectra and also show how it can be used for supervised separation of speech mixtures.
 In terms of separation performance, the ability of the network to learn cross-frame characteristics from input spectrograms enables these models to significantly outperform their feed-forward counterpart.
@@
 \label{sec:intro}
-Several neural network architectures have been proposed to develop algorithms for supervised source separation and speech enhancement~\cite{chandna2017monoaural,grais2017single, venkataramani2017end, park2016fully}.
+
+% Several neural network architectures have been proposed to develop algorithms for supervised source separation and speech enhancement~\cite{chandna2017monoaural,grais2017single, venkataramani2017end, park2016fully}.
 Currently, these networks are trained to learn discriminative audio models extensively.
@@
 
-A popular technique to learn transferable audio models for supervised source separation is the use of Non-negative matrix factorization~(NMF).
-Non-negative matrix factorization~(NMF) matrix of non-negative elements~$\mathbf{X}\in \mathbb{R}_{M\times N}^{\geq0}$ as a product of the basis matrix~$\mathbf{W}$ and the activation matrix~$\mathbf{H}$.
+Non-negative matrix factorization~(NMF) of magnitude-spectrograms has been a very popular method for modeling sources for supervised source separation applications~\cite{smaragdis2014static, virtanen2015compositional}.
+NMF factorizes a matrix of non-negative elements~$\mathbf{X}\in \mathbb{R}_{M\times N}^{\geq0}$ as a product of the basis matrix~$\mathbf{W}$ and the activation matrix~$\mathbf{H}$.
 The notation $\mathbb{R}_{M \times N}^{\geq0}$ represents the set of matrices of non-negative elements of size $M \times N$.
@@
 Additionally, this interpretation enables a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance.
-\\
 
@@
 However, NMF and its neural network equivalent are unable to explicitly utilize these cross-frame patterns available in a spectrogram.
-To alleviate this drawback, Smaragdis~\cite{smaragdis2007convolutive} proposed a convolutive version to NMF that allows spectro-temporal patterns as representative basis elements.
+To alleviate this drawback, Smaragdis~\cite{smaragdis2007convolutive} proposed a convolutive version to NMF~(conv-NMF) that allows spectro-temporal patterns as representative basis elements.
 In this paper, we develop a neural network alternative to such convolutive audio models for supervised source separation.
@@
 Utilizing the models to separate a source from a given mixture forms the second step.
+
+Developing a neural network alternative to conv-NMF also results in a significant advantage in terms of model re-usability.
+Several neural network architectures have been recently proposed for supervised source separation~\cite{grais2017single, venkataramani2017end}.
+These networks are trained to learn discriminative audio models by training them on mixture spectrograms.
+Thus, the networks learn a decomposition that works only for a specific source-interference pair.
+Developing an alternative to conv-NMF allows us to learn audio-models by training neural networks on clean source spectrograms.
+These models can be easily re-used irrespective of the interfering component.
 The remainder of the paper is organized as follows.
@@
 We evaluate these models in terms of their separation performance in section~\ref{sec:} and conclude in section~\ref{cite:}.
-\\
 
@@
 The notation $\mathbf{X}(i,j)$ represents the element of $\mathbf{X}$ indexed by the $i^{\text{th}}$ row and the $j^{\text{th}}$ column.
-Thus, we can interpret this operation as a two-step convolutional auto-encoder~(CAE) as follows,
+We can interpret this operation as a two-layer convolutional auto-encoder~(CAE) as follows,
 
 \begin{align}
-    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= \sum_{i=1}^{r} \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k) \\
+    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k) \\
     \text{$2^{\text{nd}}$ layer:~}\hat{\mathbf{X}}(f,t) &= \sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)
@@
 subject to non-negativity of $\mathbf{W}_{i}$ and $\mathbf{H}$.
-Here, we assume that the convolutional filters~$\mathbf{W},~\mathbf{W}^{\ddagger}$ have a size of $M\times T$ where, $T$ represents the depth of the convolution.
+Here, we assume that the convolutional layer filters~$\mathbf{W},~\mathbf{W}^{\ddagger}$ have a size of $M\times T$ where, $T$ represents the depth of the convolution and $M$ denotes the height of the input matrix~$\mathbf{X}$.
 In this representation, $\mathbf{W}_{i}$ and $\mathbf{H}$ correspond to the $i^{\text{th}}$ basis matrix and the activation matrix respectively.
 The filters of the first convolutional neural network~(CNN) act as inverse filters in defining the auto-encoder.
-In the remainder of this section, we will refer to the first convolutional layer as the encoder that estimates a code from the input representation.
-The second CNN layer generates an approximation of the input from the code and will be referred to as the decoder.
+In the remainder of this section, we will refer to the first convolutional layer as the ``encoder'' that estimates a code from the input representation.
+The second CNN layer generates an approximation of the input from the code and will be referred to as the ``decoder''.
 We can simplify the non-negativity constraints by incorporating a non-linearity into the definitions of the encoder and the decoder.
@@
 \begin{align}
-    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= g\left(\sum_{i=1}^{r} \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k)\right) \\
+    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= g\left(\sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k)\right) \\
     \text{$2^{\text{nd}}$ layer:~}\hat{\mathbf{X}}(f,t) &= g\left(\sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)\right)
@@
 Here, the non-linearity $g(.):\mathbb{R}\rightarrow \mathbb{R}^{\geq0}$ applies an element-wise non-linearity and ensures that the activation matrix and the reconstruction are non-negative.
-We now note a couple of key points about the CAE.
+In this paper, we use the rectified-linear unit as the non-linearity.
+We now note some key points about the CAE.
 (i) The output of the encoder gives the latent representation of the decomposition.
@@
 Thus, the basis matrices (decoder filters) can assume negative values.
-
-To train the auto-encoder and learn the basis and activation matrices, in the remainder of the paper, we apply the following approach.
-We minimize the KL-divergence between the input spectrogram~$\mathbf{X}$ and its reconstruction~$\hat{\mathbf{X}}$ given by,
+To train the auto-encoder, we minimize the KL-divergence between the input spectrogram~$\mathbf{X}$ and its reconstruction~$\hat{\mathbf{X}}$ given by,
 \begin{equation*}
@@
 \label{subsec:practical}
-Having developed the CAE equivalent to convolutive NMF, we can now begin to understand the nature of the basis and activation matrices learned by the network.
+Having developed the CAE equivalent to conv-NMF, we can now begin to understand the nature of the basis and activation matrices learned by the network.
 To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{cnn_demo}.
@@
 \centering
-  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\linewidth]{Figs/SeparationPerformance.png}
-  \caption{...}~\label{fig:separation_performance}
+  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=0.95\linewidth]{Figs/CNN_Separation_Performance.png}
+  \caption{Separation performance of convolutive models obtained using a CNN-CNN Auto-encoder for varying values of $K$. }~\label{fig:cnnseparation_performance}
 \end{figure*}
@@
 The neural networks are initialized using the Xavier initialization scheme~\cite{glorot2010understanding}.
-The networks are trained by applying a batch gradient descent training procedure with a batchsize of $18$ frames and the parameters updated using the RMSProp algorithm~\cite{tieleman2012rmsprop}, with a learning rate and momentum of $0.001$ and $0.7$ respectively.
+The networks are trained by applying a batch gradient descent training procedure and the parameters updated using the RMSProp algorithm~\cite{tieleman2012rmsprop}, with a learning rate and momentum of $0.001$ and $0.7$ respectively.
 
@@
 \label{subsec:results}
-Figure~\ref{fig:separation_performance} gives the separation performance for the CNN-CNN models (left) and RNN-CNN models (right) for varying values of number of filters~$K$.
+Figure~\ref{fig:cnnseparation_performance} gives the separation performance for the CNN-CNN models (left) and RNN-CNN models (right) for varying values of number of filters~$K$.
 We evaluate the proposed CAE models by comparing the separation performance to their equivalent feed-forward~(FF) counterparts proposed in~\cite{smaragdis2017aneural}.

### 2017-06-11 paper/paper.tex
I am pushing my additions on paper (into github)

@@
 The second CNN layer generates an approximation of the input from the code and will be referred to as the ``decoder''.
-We can simplify the non-negativity constraints by incorporating a non-linearity into the definitions of the encoder and the decoder.
+We can satisfy the non-negativity constraints by incorporating a non-linearity into the definitions of the encoder and the decoder.
 Thus,
@@
 Here, the non-linearity $g(.):\mathbb{R}\rightarrow \mathbb{R}^{\geq0}$ applies an element-wise non-linearity and ensures that the activation matrix and the reconstruction are non-negative.
-In this paper, we use the rectified-linear unit as the non-linearity.
+In our experiments, we used the soft-plus function which is given by the formula $SP(x) = \log(1+\exp(x))$ as the non-linearity.
 We now note some key points about the CAE.
-(i) The output of the encoder gives the latent representation of the decomposition.
+(i) The output of the encoder gives the latent representation (analog of activations in NMF \textbf{cem:we do not give the reference to corresponding equation number here} of the decomposition.
 (ii) The weights of the decoder act as the basis vectors of the decomposition.
@@
 
-\section{Extensions to infinite summation}
+\section{Using an Infinite Impulse Response (IIR) filter in the encoder}
 \label{sec:rnncnn}
+
+As we have discussed above, the encoder part acts as the inverse filter to the decoder.
+We can think of a convolutional networks as a filter with a finite impulse response (FIR).
+As known from DSP theory, the inverse of an FIR filter is an IIR filter.
+Therefore, we explore using a Recurrent Neural Network to get an IIR encoder.
+The encoder equation is obtained as follows:
+
+\begin{align}
+  \mathbf Z(k_1, t, i) =& g\Bigg( \sum_{k_2}\mathbf W^i(k_1,k_2 ) \mathbf Z(k_2, t-1, i) + \notag \\
+  & \hspace{2cm} \sum_l \mathbf U^i(k_1,l) \mathbf X(l, t) \Bigg), \\
+  \mathbf H(i, t) =& \sum_{k_1} \mathbf Z^i(k_1, t)
+\end{align}
+where
 

### 2017-06-11 main.py
I have made some improments for the toy example part

@@ -101,6 +101,8 @@ def model_driver(d,data):
             all_times, tr_logls, _, _ = rnn1.optimizer(data = data, 
                     rnn_handles = rnn1_handles, sess = sess)
 
+            
+
 
 def main(dictionary):
     #set the model categories
@@ -199,7 +201,7 @@ def main(dictionary):
 #import matplotlib.pyplot as plt
 wform = 'block_diagonal'# either diagonal, block_diagonal or full
 input_dictionary = {'seedin' : [1144, 1521], #setting the random seed. First is for numpy, second is for tensorflow 
-            'task' : 'source_sep', #this helps us how to load the data with the load_data function in rnns.py 
+            'task' : 'toy_example', #this helps us how to load the data with the load_data function in rnns.py 
             'data' : 'timit', #the dataset, options are inside the load_data function 
             'encoder': 'mb_mod_lstm', #options are: feed_forward, convolutive, mb_mod_lstm
             'decoder': 'convolutive',
@@ -207,9 +209,9 @@ input_dictionary = {'seedin' : [1144, 1521], #setting the random seed. First is 
             'wform_global' : wform,
             'num_configs' : 60, #number of hyper parameter configurations to be tried 
             'start' : 0,  #this is used to start from a certain point (can be useful with fixed seed, or when hyper-parameters are loaded) 
-            'EP' : 2000, #number of epochs per run 
+            'EP' : 200, #number of epochs per run 
             'dropout' : [1, 1], #first is the input second is the output keep probability 
-            'device' : 'gpu:0', #the device to be used in the computations 
+            'device' : 'gpu:1', #the device to be used in the computations 
             'server': socket.gethostname(),
             'verbose': True, #this prints out the batch location
             'load_hparams': True, #this loads hyper-parameters from a results file
@@ -221,6 +223,7 @@ input_dictionary = {'seedin' : [1144, 1521], #setting the random seed. First is 
             'optimizer':'RMSProp', #options are, Adam, RMSProp, Adadelta
             'activation':'softplus',
             'separation_method':'complete',
+            'sparsity_coef': 0,   # sparsity tradeoff parameter (only used for toy examples)
             'notes':''} 
 
 perfs = main(input_dictionary)

### 2017-06-11 rnns.py

@@ -9,7 +9,7 @@ import sklearn.feature_extraction.text as ft
 import itertools
 import sys
 import os
-#import matplotlib.pyplot as plt
+import matplotlib.pyplot as plt
 import scipy.io.wavfile as wavfile
 from music_utilities import *
 
@@ -285,6 +285,8 @@ class rnn(object):
             #define the cost         
             eps = 1e-16
             cost = tf.reduce_mean( y*(tf.log( y + eps ) - tf.log( yhat + eps ))  - y + yhat ) 
+            if d['task'] == 'toy_example':
+                cost = cost + d['sparsity_coef']*tf.reduce_sum(hhat)
 
             #define the optimizer
             if d['optimizer'] == 'Adam':
@@ -314,6 +316,9 @@ class rnn(object):
                              'relevant_inds':relevant_inds,
                              'targets':targets,
                              }
+
+            if d['task'] == 'toy_example':
+                graph_handles.update({'all_variables': tvars})
                                            
             return graph_handles
 
@@ -634,17 +639,26 @@ class rnn(object):
 
         Hhat, Yhat = sess.run([rnn_handles['h'], rnn_handles['preds']], feed) 
 
+        pdb.set_trace()
+        if d['task' ] == 'toy_example':
+            all_variables = rnn_handles['all_variables']
+            
+            for var in all_variables:
+                if 'decoder/fltr' in var.name:
+                    filter_weights = sess.run(var, feed).squeeze()
+                    filter_weights = np.transpose(filter_weights, [2, 0, 1])
+
+        pdb.set_trace()
+
         return all_times, tr_logls, test_logls, valid_logls
 
     def input_optimizer(self, data, rnn_handles, sess):
         d = self.model_specs # unpack the variables 
     
         tr = SimpleDataIterator(data, num_buckets = d['num_buckets'])
-        #tst = SimpleDataIterator(data['Test'])
-        #valid = SimpleDataIterator(data['Validation'])
-
+        
         all_times, tr_logls, test_logls, valid_logls = [], [], [], [] 
-        for ep in range(2*d['EP']):
+        for ep in range(d['EP']):
             t1, tr_logl = time.time(), []
             while tr.epochs == ep:
                 trb = tr.next_batch(
@@ -667,15 +681,6 @@ class rnn(object):
         
[…]

### 2017-06-11 paper/paper.tex
Added speech bases figure and updated some text

@@
 
+\begin{figure}[t]
+\centering
+  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\linewidth]{Figs/CNN_demo.png}
+  \caption{...}~\label{fig:cnn_demo}
+\end{figure}
+
 \section{Non-negative Convolutiional Auto-encoders}
@@
 \end{align}
-Here, the non-linearity $g(.):\mathbb{R}\rightarrow \mathbb{R}^{\geq0}$ applies an element-wise non-linearity and ensures that the activation matrix and the reconstruction are non-negative.
+Here, the $g(.):\mathbb{R}\rightarrow \mathbb{R}^{\geq0}$ applies an element-wise non-linearity and ensures that the activation matrix and the reconstruction are non-negative.
 In our experiments, we used the soft-plus function which is given by the formula $SP(x) = \log(1+\exp(x))$ as the non-linearity.
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 \centering
-  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\linewidth]{Figs/CNN_demo.png}
-  \caption{...}~\label{fig:cnn_demo}
-\end{figure}
+  \includegraphics[clip, trim = 3cm 2.4cm 3cm 2.1cm, width=\linewidth, height = 0.7\columnwidth]{Figs/Bases.eps}
+  \caption{Decoder filters obtained by training the CAE on magnitude-spectrograms of utterances of a male speaker.
+This decomposition is obtained for the configuration $K = 80$ and $T = 8$.
+We see that the filters resemble snippets of a speech spectrogram.}~\label{fig:cnn_demo_speech}
+\end{figure*}
 
@@
 We also incorporate sparsity constraints on the activation, i.e., the output of the first CNN.
-We see that the CNN filter resembles a snippet of the input spectrogram.
-From the nature of the activation, it is clear that the encoder acts as a matched-filter and identifies the points in time when the corresponding pattern becomes active.
+In this example, the CNN filter resembles a snippet of the input spectrogram.
+We also see that the activation comprises a series of impulse trains.
+Thus, the encoder acts as a matched-filter and identifies the points in time when the corresponding pattern becomes active.
 As shown in (\ref{eq:nmfcnncnn}), the time-frequency pattern is captured by the filters of the decoder.
-Thus, the encoder learns the inverse-filter of the decoder.
+Given the nature of the activation we see that the encoder attempts to learn the inverse filter to the decoder.
+Figure~\ref{fig:cnn_demo_speech} shows the decoder filters obtained by training the CAE on speech utterances of a male speaker.
+Similar to the previous toy-example, the decoder filters learn patterns that resemble snippets of a speech spectrogram.
 
-% The approximation $\hat X$ for a given spectrogram $X$ is computed as follows:
-
-% \begin{align}
-% 	\hat H(k,t) =& \sigma_1\left ( \sum_{f,t'} X(f,t-t') F_e(f,t',k) \right ) \notag \\
-% 	\hat X(f,t) =& \sigma_2 \left ( \sum_k \sum_{t'} \hat H(k,t-t') F_d(f, t',k) \right )
-% \end{align}
+The decoder CNN can be interpreted as a finite impulse response~(FIR) filter.
+The encoder attempts to approximate the inverse of the decoder FIR filter as an FIR filter.
+From our knowledge of signal processing, it can be shown that the inverse filter of an FIR filter has an infinite impulse response~(IIR)~\cite{} (**Give reference**).
+Thus, approximating the encoder as an FIR filter or a CNN will hinder the decoder from learning patterns that are invertible only by an IIR filter.
+One such example is shown in figure~\ref{fig:}.
+Therefore, we explore using a Recurrent Neural Network~(RNN) to get an IIR encoder.
 
@@
 
-As we have discussed above, the encoder part acts as the inverse filter to the decoder.
-We can think of a convolutional networks as a filter with a finite impulse response (FIR).
+As discussed above, the encoder part acts as the inverse filter to the decoder.
+We can think of a CNN as a filter with a finite impulse response.
 As known from DSP theory, the inverse of an FIR filter is an IIR filter.
-Therefore, we explore using a Recurrent Neural Network to get an IIR encoder.
-The encoder equation is obtained as follows:
+To the best of our knowledge, this results in a novel RNN-CNN auto-encoder architecture.
+In the remainder of the paper, we will refer to this auto-encoder as a Recurrent-Convolutive Auto-encoder~(RCAE).
+The encoder RNN equation is obtained as follows:
 

### 2017-06-12 paper/paper.tex

@@
 \usepackage{amsfonts}
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows,positioning,calc}
 
@@
 %
+
+\tikzset{
+block/.style = {draw, fill=white, rectangle, minimum height=3em, minimum width=3em},
+tmp/.style  = {coordinate}, 
+sum/.style= {draw, fill=white, circle, node distance=1cm},
+input/.style = {coordinate},
+output/.style= {coordinate},
+pinstyle/.style = {pin edge={to-,thin,black}
+}
+}
+
 \begin{document}
@@
 
+\begin{figure*}
+  \centering
+  \begin{tikzpicture}[auto, node distance=2cm,>=latex']
+
+      \node [,name=input] (input) {$X$};
+      \coordinate [, name=rinput, right of=input, node distance=2cm ](rinput) {};
+
+      \node [, right of=rinput] (controller) {$\vdots$};
+      \node [block, above of=controller,node distance=1.3cm] (up){$\mathbf W^\ddagger_1$};
+      \node [block, below of=controller,node distance=1.3cm] (rate) {$\mathbf W^\ddagger_K$};
+      \node [sum, right of=controller,node distance=2cm] (sum2) {CT};
+
+      \node [block, right of=sum2,node distance=2cm] (system) {$\frac{a_{\beta 2}}{s+a_{\beta 1}}$};
+      \node [output, right of=system, node distance=2cm] (output) {};
+      \node [tmp, below of=controller] (tmp1){$H(s)$};
+
+      \draw [->] (rinput) -- (controller); 
+      \draw [->] (controller) -- (sum2);
+      \draw [->] (sum2) -- node{$U(s)$} (system);
+     % \draw [->] (system) -- node [name=y] {$Y(s)$}(output);
+      \draw [->] (rinput) |- (rate);
+      \draw [->] (rate) -| (sum2);
+      \draw [->] (rinput) |- (up);
+      \draw [->] (up) -| (sum2);
+      % \draw [->] (y) |- (tmp1)-| node[pos=0.99] {$-$} (sum1);
+      %\draw [->] (extra)--(sum2);
+      %\draw [->] ($(0,1.5cm)+(extra)$)node[above]{$d_{\beta 2}$} -- (extra);
+    \end{tikzpicture}
+  \caption{Block Diagram of CNN-CNN Autoencoder} 
+  \label{CNN-CNN Auto Encoder}
+\end{figure*}
+
 \subsection{Practical Considerations}
@@
 
-As discussed above, the encoder part acts as the inverse filter to the decoder.
+%As discussed above, the encoder part acts as the inverse filter to the decoder.
 We can think of a CNN as a filter with a finite impulse response.
@@
 In the remainder of the paper, we will refer to this auto-encoder as a Recurrent-Convolutive Auto-encoder~(RCAE).
-The encoder RNN equation is obtained as follows:
+
+In this section, the goal is to construct a recurrent encoder analogous to the convolutive encoder we discussed in the previous section.
+The potential gain of using a recurrent encoder over a finite length convolutive encoder is due to the fact that a recurrent filter in theory can capture arbitrarily long temporal dependencies.
+From a signal processing point of view, the motivation for going with a recurrent filter is that, the inverse of an finite length filter (the convolutive basis in the encoder) is given by a recurrent filter.
+We explore the idea, and obtain source separation results on the same dataset.
+
+The way we go about building an encoder is by passing the input through $K$ separate recurrent neural networks (RNNs) (Note that $K$ was the number of filters/components in the convolutive model).
+The $k$'th RNN recursion in the encoder is given by the following equation:
 
 \begin{align}
-  \mathbf Z(k_1, t, i) =& g\Bigg( \sum_{k_2}\mathbf W^i(k_1,k_2 ) \mathbf Z(k_2, t-1, i) + \notag \\
-  & \hspace{2cm} \sum_l \mathbf U^i(k_1,l) \mathbf X(l, t) \Bigg), \\
-  \mathbf H(i, t) =& \sum_{k_1} \mathbf Z^i(k_1, t)
+  \mathbf Z(k_1, t, k) =& \tanh \Bigg( \sum_{k_2 = 1}^{K_{in}} \mathbf W^k(k_1,k_2 ) \mathbf Z(k_2, t-1, k) + \notag \\
+  & \hspace{0cm} \sum_l \mathbf U^k(k_1,l) \mathbf X(l, t) \Bigg),\; k \in \{1,\dots,K\}, 
 \end{align}
-where
+where $\mathbf Z(:, t, k) \in \mathbb R^{K_{in}}$ denotes the latent state vector of the $k$'th RNN at time $t$.
+We denote the hidden state dimensionality of each RNN with $K_{in}$.
+The recurrent and projection matrices of the $k'th$ RNN are respectively denoted with $\mathbf W^k$, and $\mathbf U^k$.
+Note that although the given recursion corresponds to the vanilla-RNN architecture, there is no restriction on the RNN architecture choice.
+In our experiments, we have used the LSTM architecture [cite].
+After going through the RNN recursions, the encoder output $\mathbf H(i, t)$ is obtained by summing the RNN outputs over the first dimension:
+\begin{align}
+  \mathbf H(i, t) = \sum_{k_1=1}^{K_{in}} \mathbf Z(k_1, t, i)
+\end{align}
 

### 2017-06-12 paper/paper.tex
Updated results section and figs

--- 2017-06-11 paper/paper.tex
@@
 
-\begin{figure*}[t]
+\begin{figure}[ht!]
 \centering
-  \includegraphics[clip, trim = 3cm 2.4cm 3cm 2.1cm, width=\linewidth, height = 0.7\columnwidth]{Figs/Bases.eps}
-  \caption{Decoder filters obtained by training the CAE on magnitude-spectrograms of utterances of a male speaker.
+  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\columnwidth, height = 0.8\columnwidth]{Figs/Bases.png}
+  \caption{A subset of decoder filters obtained by training the CAE on magnitude-spectrograms of utterances of a male speaker.
 This decomposition is obtained for the configuration $K = 80$ and $T = 8$.
 We see that the filters resemble snippets of a speech spectrogram.}~\label{fig:cnn_demo_speech}
-\end{figure*}
+\end{figure}
 
@@
 \centering
-  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=0.95\linewidth]{Figs/CNN_Separation_Performance.png}
+  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\linewidth]{Figs/FullPerformance.png}
   \caption{Separation performance of convolutive models obtained using a CNN-CNN Auto-encoder for varying values of $K$. }~\label{fig:cnnseparation_performance}
@@
 Although the SAR for CAE models degrades slightly as compared to FF models, the improvement in significant improvement in SIR compensates for this loss.
+The performance of the models achieves a peak value for $K=80$.
+However, the separation performance does not degrade significantly for other values of $K$.
+Thus, the choice of $K$ does not appear to be a very critical consideration for these models.
+
+The convolutive speech models obtained by training the RCAE also outperform the FF auto-encoder models, as seen by the median values and inter-quartile ranges.
+This improvement in performance is not as pronounced as the CAE models for the case of speech mixtures.
+However, the experimentation serves as a proof of concept that exploring non-uniform auto-encoder architectures could lead to interesting and potentially powerful algorithms for several applications.
 

### 2017-06-12 paper/paper.tex
Added the CNN-CNN block diagram

--- 2017-06-12 paper/paper.tex
@@
 
-      \node [,name=input] (input) {$X$};
-      \coordinate [, name=rinput, right of=input, node distance=2cm ](rinput) {};
+      \node [,name=input] (input) {$\mathbf X$};
+      \coordinate [, name=rinput, right of=input, node distance=1.3cm ](rinput) {};
+      \node [block, right of=rinput, node distance=1.3cm] (controller) {$\vdots$};
+      \node [sum, right of=controller, node distance=2cm] (controllers){$\sum_j$};
+      \node [block, above of=controller,node distance=1.3cm] (up){$\mathbf X *\mathbf W^\ddagger_1$};
+      \node [sum, right of=up, node distance=2cm] (ups) {$\sum_j$};
+      \node [block, below of=controller,node distance=1.3cm] (rate) {$\mathbf X*\mathbf W^\ddagger_K$};
+      \node [sum, right of=rate, node distance=2cm] (rates) {$\sum_j$};
 
-      \node [, right of=rinput] (controller) {$\vdots$};
-      \node [block, above of=controller,node distance=1.3cm] (up){$\mathbf W^\ddagger_1$};
-      \node [block, below of=controller,node distance=1.3cm] (rate) {$\mathbf W^\ddagger_K$};
-      \node [sum, right of=controller,node distance=2cm] (sum2) {CT};
+      \node [sum, right of=controllers,node distance=1.3cm] (sum2) {CT};
+      \node[sum, right of=sum2,node distance=2cm] (system) {$g()$};
+      \node [block, right of=system, node distance=2cm] (dec2) {$\vdots$};
+      \node [block, above of=dec2, node distance=1.3cm] (dec1) {$\mathbf H(1, :) * \mathbf W_1$};
+      \node [block, below of=dec2, node distance=1.3cm] (dec3) {$\mathbf H(K, :) *\mathbf W_K$};
+      \node [sum, right of=dec2, node distance=2cm] (sum3) {$+$};
+      \node [sum, right of=sum3, node distance=1.3cm] (g2) {$g()$};
+      \node [,right of=g2, node distance=1.3cm] (outp) {$\widehat{\mathbf X}$};
 
-      \node [block, right of=sum2,node distance=2cm] (system) {$\frac{a_{\beta 2}}{s+a_{\beta 1}}$};
-      \node [output, right of=system, node distance=2cm] (output) {};
-      \node [tmp, below of=controller] (tmp1){$H(s)$};
-
+      \draw [-] (input) -- (rinput);
       \draw [->] (rinput) -- (controller); 
-      \draw [->] (controller) -- (sum2);
-      \draw [->] (sum2) -- node{$U(s)$} (system);
-     % \draw [->] (system) -- node [name=y] {$Y(s)$}(output);
+      \draw [->] (controller) -- (controllers);
+      \draw [->] (controllers) -- (sum2);
+      \draw [->] (sum2) -- (system);
       \draw [->] (rinput) |- (rate);
-      \draw [->] (rate) -| (sum2);
+      \draw [->] (rate) -- (rates);
+      \draw [->] (rates) -| (sum2);
       \draw [->] (rinput) |- (up);
-      \draw [->] (up) -| (sum2);
-      % \draw [->] (y) |- (tmp1)-| node[pos=0.99] {$-$} (sum1);
-      %\draw [->] (extra)--(sum2);
-      %\draw [->] ($(0,1.5cm)+(extra)$)node[above]{$d_{\beta 2}$} -- (extra);
+      \draw [->] (up) -- (ups);
+      \draw [->] (ups) -| (sum2);
+      \draw [->] (system) -- node {$\mathbf H(.,:)$} (dec2);
+      \draw [->] (system) |- node {$\mathbf H(1,:)$} (dec1);
+      \draw [->] (system) |- node[anchor=north]{$\mathbf H(K,:)$} (dec3);
+      \draw [->] (dec1) -| (sum3);
+      \draw [->] (dec2) -- (sum3);
+      \draw [->] (dec3) -| (sum3);
+      \draw [->] (sum3) -- (g2);
+      \draw [->] (g2) -- (outp);
     \end{tikzpicture}

### 2017-06-12 paper/paper.tex
Conclusion added

@@
   \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\linewidth]{Figs/CNN_demo.png}
-  \caption{...}~\label{fig:cnn_demo}
+  \caption{Basis decomposition of a toy-illustration obtained using a CNN-CNN auto-encoder.}~\label{fig:cnn_demo}
 \end{figure}
@@
 In our experiments, we used the soft-plus function which is given by the formula $SP(x) = \log(1+\exp(x))$ as the non-linearity.
-We now note some key points about the CAE.
-(i) The output of the encoder gives the latent representation (analog of activations in NMF \textbf{cem:we do not give the reference to corresponding equation number here} of the decomposition.
+Based on (\ref{eq:nmfcnncnn}), we now note some key points about the CAE.
+(i) The output of the encoder gives the latent representation (analog of activations in NMF) of the decomposition.
 (ii) The weights of the decoder act as the basis vectors of the decomposition.
@@
 \centering
-  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\columnwidth, height = 0.8\columnwidth]{Figs/Bases.png}
+  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\columnwidth, height = \columnwidth]{Figs/Bases.png}
   \caption{A subset of decoder filters obtained by training the CAE on magnitude-spectrograms of utterances of a male speaker.
@@
 \end{figure}
+Although the filters can assume negative values, the use of a non-linearity does not allow cross-cancellations across the basis elements.
 
@@
 \label{sec:conclusion}
+In this paper, we developed and investigated the use of a convolutional auto-encoder as an alternative to learn convolutive basis decompositions of speech and audio signals.
+The ability of the networks to include temporal dependencies allow the auto-encoders to learn cross-frame structures in the input spectrogram.
+We demonstrated that this results in a significant improvement in separation performance as compared to feed-forward auto-encoder models.
+This approach also allows for several extensions and generalizations to convolutive audio models, that can be easily implemented using the available diversity of neural networks.
+One such extension considered in this paper is the use of auto-encoders formed by a cascade of recurrent and convolutive layers.
+Although these models are not as potent as CAE models for separation of speech mixtures, these extensions could lead to interesting non-uniform auto-encoder architectures for several applications.
 

### 2017-06-12 paper/paper.tex
I have added the RNN encoder diagram

--- 2017-06-12 paper/paper.tex
@@
 
-\section{Non-negative Convolutiional Auto-encoders}
+\section{Non-negative Convolutional Auto-encoders}
 \subsection{Network Architecture}
@@
 \begin{align}
-    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k) \\
-    \text{$2^{\text{nd}}$ layer:~}\hat{\mathbf{X}}(f,t) &= \sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)
+  \text{$1^{st}$ layer:~}\mathbf{H}(i,t) &= \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k) \\
+  \text{$2^{nd}$ layer:~}\hat{\mathbf{X}}(f,t) &= \sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)
     \label{eq:nmfcnncnn}
@@
 Here, the $g(.):\mathbb{R}\rightarrow \mathbb{R}^{\geq0}$ applies an element-wise non-linearity and ensures that the activation matrix and the reconstruction are non-negative.
-In our experiments, we used the soft-plus function which is given by the formula $SP(x) = \log(1+\exp(x))$ as the non-linearity.
+The block diagram of the whole CNN-CNN encoder is given in Figure \ref{diag:cnncnn}.
+In our experiments, we used the soft-plus function which is given by the formula $g(x) = \log(1+\exp(x))$ as the non-linearity.
 We now note some key points about the CAE.
@@
 
-\begin{figure}[ht!]
-\centering
-  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\columnwidth, height = 0.8\columnwidth]{Figs/Bases.png}
-  \caption{A subset of decoder filters obtained by training the CAE on magnitude-spectrograms of utterances of a male speaker.
-This decomposition is obtained for the configuration $K = 80$ and $T = 8$.
-We see that the filters resemble snippets of a speech spectrogram.}~\label{fig:cnn_demo_speech}
-\end{figure}
-
-\begin{figure*}
+\begin{figure*} [t]
   \centering
@@
 
-      \node [sum, right of=controllers,node distance=1.3cm] (sum2) {CT};
-      \node[sum, right of=sum2,node distance=2cm] (system) {$g()$};
+      \node [sum, right of=controllers,node distance=1.3cm] (sum2) {$g()$};
+      \coordinate[, right of=sum2,node distance=2cm] (system) {};
       \node [block, right of=system, node distance=2cm] (dec2) {$\vdots$};
@@
       \draw [->] (controllers) -- (sum2);
-      \draw [->] (sum2) -- (system);
+      \draw [-] (sum2) --node{$\mathbf H$} (system);
       \draw [->] (rinput) |- (rate);
@@
       \draw [->] (ups) -| (sum2);
-      \draw [->] (system) -- node {$\mathbf H(.,:)$} (dec2);
+      \draw [->] (system) -- node {} (dec2);
       \draw [->] (system) |- node {$\mathbf H(1,:)$} (dec1);
@@
   \caption{Block Diagram of CNN-CNN Autoencoder} 
-  \label{CNN-CNN Auto Encoder}
+  \label{diag:cnncnn}
 \end{figure*}
+
+\begin{figure}[ht!]
+\centering
+  \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\columnwidth, height = 0.8\columnwidth]{Figs/Bases.png}
+  \caption{A subset of decoder filters obtained by training the CAE on magnitude-spectrograms of utterances of a male speaker.
+This decomposition is obtained for the configuration $K = 80$ and $T = 8$.
+We see that the filters resemble snippets of a speech spectrogram.}~\label{fig:cnn_demo_speech}
+\end{figure}
 
@@
 
-The decoder CNN can be interpreted as a finite impulse response~(FIR) filter.
+\textbf{Cem: Impulse response concept is for linear systems.
+I am not too sure about using it in this context.} The decoder CNN can be interpreted as a finite impulse response~(FIR) filter.
 The encoder attempts to approximate the inverse of the decoder FIR filter as an FIR filter.
@@
 
-\section{Using an Infinite Impulse Response (IIR) filter in the encoder}
+\section{Using a recurrent filter in the encoder}
 \label{sec:rnncnn}
@@
 \end{align}
+
+The recurrent encoder's block diagram is given in Figure \ref{diag:rnnenc}.
+
+\begin{figure*} [t]
+  \centering
+  \begin{tikzpicture}[auto, node distance=2cm,>=latex']
+
+      \node [,name=input] (input) {$\mathbf X$};
+      \coordinate [, name=rinput, right of=input, node distance=1.3cm ](rinput) {};
+      \node [block, right of=rinput, node distance=1.3cm] (controller) {$\vdots$};
+      \node [sum, right of=controller, node distance=2cm] (controllers){$\sum_j$};
+      \node [block, above of=controller,node distance=1.3cm] (up){$\text{RNN}_1(\mathbf X)$};
+      \node [sum, right of=up, node distance=2cm] (ups) {$\sum_j$};
+      \node [block, below of=controller,node distance=1.3cm] (rate) {$\text{RNN}_K(\mathbf X)$};
+      \node [sum, right of=rate, node distance=2cm] (rates) {$\sum_j$};
+
+      \node [sum, right of=controllers,node distance=1.3cm] (sum2) {$g()$};
+      \node[, right of=sum2,node distance=1.3cm] (system) {$\mathbf H$};
+
+      \draw [-] (input) -- (rinput);
+      \draw [->] (rinput) -- (controller); 
+      \draw [->] (controller) -- (controllers);
+      \draw [->] (controllers) -- (sum2);
+      \draw [->] (sum2) --(system);
+      \draw [->] (rinput) |- (rate);
+      \draw [->] (rate) -- (rates);
+      \draw [->] (rates) -| (sum2);
+      \draw [->] (rinput) |- (up);
+      \draw [->] (up) -- (ups);
+      \draw [->] (ups) -| (sum2);
+      \end{tikzpicture}
+  \caption{Block Diagram of RNN Encoder} 
+  \label{diag:rnnenc}
+\end{figure*}
 

### 2017-06-12 paper/paper.tex

@@
 \end{align}
+
 Here, the $g(.):\mathbb{R}\rightarrow \mathbb{R}^{\geq0}$ applies an element-wise non-linearity and ensures that the activation matrix and the reconstruction are non-negative.
@@
 
-\begin{figure*} [t]
+\begin{figure*}[t]
   \centering

### 2017-06-13 paper/paper.tex
Paris cursory changes, Refs updated
Paris said, he will read it tonight.

@@
 \documentclass{article}
-\usepackage{amsmath,graphicx,mlspconf}
+\usepackage{amsmath,graphicx,mlspconf,hyperref}
 \usepackage{amsfonts}
@@
 More importantly, we can now exploit the available diversity of neural networks to propose novel and potentially powerful extensions to existing auto-encoder architectures, based on a cascade of recurrent and convolutional neural network layers.
+\\
 
@@
 The rows of $\mathbf{H}$ indicate the activity of these basis vectors in time.
-\textbf{Cem: I am not sure how smooth the transition to neural nets here.} As shown in~~\cite{smaragdis2017aneural}, the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network.
+\textbf{C: I am not sure how smooth the transition to neural nets here.} As shown in~~\cite{smaragdis2017aneural}, the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network.
 We can interpret NMF as a non-negative auto-encoder in the following manner,
 \begin{align}
-    \text{$1^{\text{st}}$ layer:~}\mathbf{H} &= g(\mathbf{W^{\ddagger}} \cdot \mathbf{X}) \\
+    \text{$1^{\text{st}}$ layer:~}\mathbf{H} &= g(\mathbf{W^{\ddagger}} \cdot \mathbf{X}) \nonumber \\
     \text{$2^{\text{nd}}$ layer:~}\mathbf{X} &= g(\mathbf{W} \cdot \mathbf{H})
@@
 Additionally, this interpretation enables a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance.
+\\
 
@@
 Utilizing the models to separate a source from a given mixture forms the second step.
+\\
 
@@
 Thus, the networks learn a decomposition that works only for a specific source-interference pair.
-Developing an alternative to conv-NMF allows us to learn audio-models by training neural networks on clean source spectrograms.
+Developing an alternative to conv-NMF allows us to learn convolutive audio-models by training neural networks on clean source spectrograms.
 These models can be easily re-used irrespective of the interfering component.
+At the same time, it is also easy to propose extensions to the models by using different and exotic neural network layers.
+\\
+
 The remainder of the paper is organized as follows.
-In section~\ref{sec:}, we develop an auto-encoder that can act as an equivalent to conv-NMF audio models.
-Section~\ref{sec:} proposes a novel approach to utilize these models for supervised source separation.
-We evaluate these models in terms of their separation performance in section~\ref{sec:} and conclude in section~\ref{cite:}.
+In section~\ref{sec:conv-nmf}, we develop an auto-encoder that can act as an equivalent to conv-NMF audio models.
+Section~\ref{sec:ss} proposes a novel approach to utilize these models for supervised source separation.
+We evaluate these models in terms of their separation performance in section~\ref{sec:experiments} and conclude in section~\ref{sec:conclusion}.
+\\
 
@@
 \begin{align}
-  \text{$1^{st}$ layer:~}\mathbf{H}(i,t) &= \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k) \\
+  \text{$1^{st}$ layer:~}\mathbf{H}(i,t) &= \sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k) \nonumber \\
   \text{$2^{nd}$ layer:~}\hat{\mathbf{X}}(f,t) &= \sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)
@@
 \begin{align}
-    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= g\left(\sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k)\right) \\
+    \text{$1^{\text{st}}$ layer:~}\mathbf{H}(i,t) &= g\left(\sum_{j=0}^{M-1} \sum_{k=0}^{T-1} \mathbf{W}_{i}^{\ddagger}(j,k)\mathbf{X}(j,t-k)\right) \nonumber \\
     \text{$2^{\text{nd}}$ layer:~}\hat{\mathbf{X}}(f,t) &= g\left(\sum_{i=1}^{r} \sum_{k=0}^{T-1} \mathbf{W}_{i}(k,f)\cdot\mathbf{H}(i,t-k)\right)
@@
 In our experiments, we used the soft-plus function which is given by the formula $g(x) = \log(1+\exp(x))$ as the non-linearity.
-We now note some key points about the CAE.
-(i) The output of the encoder gives the latent representation (analog of activations in NMF \textbf{cem:we do not give the reference to corresponding equation number here} of the decomposition.
+Using (\ref{eq:nmfcnncnn}), we now note some key points about the CAE.
+(i) The output of the encoder gives the latent representation (analog of activations in NMF) of the decomposition.
 (ii) The weights of the decoder act as the basis vectors of the decomposition.
@@
 \end{figure}
-Although the filters can assume negative values, the use of a non-linearity does not allow cross-cancellations across the basis elements.
+Although the filters can assume negative values, the use of a non-linearity does not allow cross-cancellations across the basis elements. \\
 
@@
 Having developed the CAE equivalent to conv-NMF, we can now begin to understand the nature of the basis and activation matrices learned by the network.
-To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{cnn_demo}.
+To do so, we train the convolutive auto-encoder defined by~(\ref{eq:nmfcnncnn}) on a simple toy example as shown in figure~\ref{fig:cnn_demo}.
 We also incorporate sparsity constraints on the activation, i.e., the output of the first CNN.
@@
 Similar to the previous toy-example, the decoder filters learn patterns that resemble snippets of a speech spectrogram.
+\\
 
-\textbf{Cem: Impulse response concept is for linear systems.
+\textbf{C: Impulse response concept is for linear systems.
 I am not too sure about using it in this context.} The decoder CNN can be interpreted as a finite impulse response~(FIR) filter.
 The encoder attempts to approximate the inverse of the decoder FIR filter as an FIR filter.
-From our knowledge of signal processing, it can be shown that the inverse filter of an FIR filter has an infinite impulse response~(IIR)~\cite{} (**Give reference**).
+From our knowledge of signal processing, it can be shown that the inverse filter of an FIR filter has an infinite impulse response~(IIR)~\cite{smith2011spectral}\footnote{\text{http://ccrma.stanford.edu/\~{}jos/fp/Inverse\_Filters.html}}.
 Thus, approximating the encoder as an FIR filter or a CNN will hinder the decoder from learning patterns that are invertible only by an IIR filter.
-One such example is shown in figure~\ref{fig:}.
 Therefore, we explore using a Recurrent Neural Network~(RNN) to get an IIR encoder.
+\\
 
@@
 \label{sec:rnncnn}
-
-%As discussed above, the encoder part acts as the inverse filter to the decoder.
-We can think of a CNN as a filter with a finite impulse response.
-As known from DSP theory, the inverse of an FIR filter is an IIR filter.
-To the best of our knowledge, this results in a novel RNN-CNN auto-encoder architecture.
-In the remainder of the paper, we will refer to this auto-encoder as a Recurrent-Convolutive Auto-encoder~(RCAE).
-
 In this section, the goal is to construct a recurrent encoder analogous to the convolutive encoder we discussed in the previous section.
+We will refer to this auto-encoder as a Recurrent-Convolutive Auto-encoder~(RCAE).
 The potential gain of using a recurrent encoder over a finite length convolutive encoder is due to the fact that a recurrent filter in theory can capture arbitrarily long temporal dependencies.
@@
 We explore the idea, and obtain source separation results on the same dataset.
+\\
 
@@
 \end{align}
+The recurrent encoder's block diagram is given in Figure \ref{diag:rnnenc}. \\
 
-The recurrent encoder's block diagram is given in Figure \ref{diag:rnnenc}.
-
-\begin{figure*}[t]
+\begin{figure}[t]
   \centering
@@
   \label{diag:rnnenc}
-\end{figure*}
-
-% This is the same as CNN-CNN case, except the computation of the activations $H$.
-In the CNN encoder, each filter was of finite length.
-With RNN-CNN version, we are attempting to use an infinite length filter.
-The computation of $\hat H$ is as follows:
-
-% \begin{align}
-% 	Z(:,k,t) =& \sigma( W Z(:,k,t-1) + U X(:,t)) \notag \\
-% 	\hat H(k,t) =& \sum_f Z(f,k,t)
-% \end{align}
-
-% Give a toy example which shows what this model can do that CNN-CNN can not.
-
-% \section{Multi-layer Extensions}
-% \label{sec:multlayer}
+\end{figure}
 
@@
 We now turn our attention to the problem of using the models for separating the source in an unknown mixture.
+\\
 
-The previous approach to source separation~\cite{smaragdis2017aneural} using neural-network based audio-models involves estimating the latent representation of the bases, which captures the contribution of the bases to each frame of the mixture spectrogram.
-This does not utilize the encoder component of the auto-encoder for the separation task.
+The previous approach to source separation using feed-forward neural-networks~\cite{smaragdis2017aneural} involves estimating the latent representation of the bases.
+The latent representation captures the contribution of the bases to each frame of the mixture spectrogram.
+In this approach, we do not utilize the encoder of the trained auto-encoder network for the separation task.
 In this paper, we present a novel-separation scheme that utilizes the complete auto-encoder architecture for separation.
 We do so by using the following setup for separation.
-
-Given an input spectrogram $\mathbf{X}$, the auto-encoder produces an approximation of the input spectrogram which is a linear combination of its weights. We will denote to this approximation as,
+Given an input spectrogram $\mathbf{X}$, the auto-encoder produces an approximation of the input spectrogram which is a linear combination of its weights.
+We will denote to this approximation as,
 \begin{equation}
@@
 To do so, we train the network defined by~(\ref{eq:separation}) for an appropriate input $\mathbf{X}_{1}$,~$\mathbf{X}_{2}$, instead of training for the weights of the network.
-As before, we minimize the KL divergence between mixture spectrogram $\mathbf{X}_{m}$ and its approximation~$\mathbf{X}_{1} + \mathbf{X}_{2}$.
+As before, we minimize the KL divergence between mixture spectrogram $\mathbf{X}_{m}$ and its approximation~$(\mathbf{X}_{1} + \mathbf{X}_{2})$.
 Conceptually, this problem is not different to training a neural network.
 The equivalence can be seen by applying a transposition to the auto-encoder definitions in (\ref{eq:nmfcnncnn}).
-This also allows a generalized separation procedure to be applied even when the underlying architectures of the auto-encoders are changed.
+This approach provides a generalized separation procedure that can be used even when the underlying suto-encoder architectures are changed.
+\\
 
@@
 Also, $\odot$ represents the element-wise multiplication operation and the division is also element-wise.
+\\
 
@@
   \includegraphics[clip, trim = 0cm 0cm 0cm 0cm, width=\linewidth]{Figs/FullPerformance.png}
-  \caption{Separation performance of convolutive models obtained using a CNN-CNN Auto-encoder for varying values of $K$. }~\label{fig:cnnseparation_performance}
+  \caption{Separation performance of convolutive models obtained using CAE~(green) and RCAE~(red) for varying values of $K$.
+We compare these models to feed-forward auto-encoder based models~(blue) in~\cite{smaragdis2017aneural}.
+The legend indicates the encoder of the corresponding architectures.
+}~\label{fig:cnnseparation_performance}
 \end{figure*}
@@
 The magnitude spectrogram is then given as an input to the network.
+\\
 
@@
 The networks are trained by applying a batch gradient descent training procedure and the parameters updated using the RMSProp algorithm~\cite{tieleman2012rmsprop}, with a learning rate and momentum of $0.001$ and $0.7$ respectively.
+\\
 
@@
 We compare the separation performance in terms of median BSS\_eval metrics~\cite{fevotte2005bss_eval} viz., signal-to-distortion (SDR), signal-to-interference (SIR) and signal-to-artifact ratio (SAR) parameters.
+\\
 
@@
 \label{subsec:results}
-Figure~\ref{fig:cnnseparation_performance} gives the separation performance for the CNN-CNN models (left) and RNN-CNN models (right) for varying values of number of filters~$K$.
-We evaluate the proposed CAE models by comparing the separation performance to their equivalent feed-forward~(FF) counterparts proposed in~\cite{smaragdis2017aneural}.
+Figure~\ref{fig:cnnseparation_performance} gives the separation performance for the CNN-CNN models (green) and RNN-CNN models (red) for varying values of number of filters~$K$.
+We evaluate the proposed CAE models by comparing the separation performance to their equivalent feed-forward~(FF) counterparts (blue) proposed in~\cite{smaragdis2017aneural}.
 In order to maintain a uniform experimental setup, we apply the separation scheme described in~\ref{sec:ss} to all the models.
-We plot the results in terms of a violin plot that also indicates the median value and the corresponding inter-quartile range for each $K$.
+We plot the results in terms of a violin plot that also indicates the median value (dotted line in the centre) and the corresponding inter-quartile range (outer dotted lines) for each $K$.
+\\
 
@@
 This can be seen from the fact that the inter-quartile range in SDR for CAE models is higher than the inter-quartile range of corresponding FF models for several values of $K$.
-This improvement is a consequence of reduced interference generated by these models in source separation (See SIR plots).
-Although the SAR for CAE models degrades slightly as compared to FF models, the improvement in significant improvement in SIR compensates for this loss.
+This improvement is a consequence of reduced interference generated by these models in source separation (as seen in the SIR plots).
+Although the SAR for CAE models degrades slightly as compared to FF models, the significant improvement in SIR compensates for this loss.
 The performance of the models achieves a peak value for $K=80$.
 However, the separation performance does not degrade significantly for other values of $K$.
-Thus, the choice of $K$ does not appear to be a very critical consideration for these models.
+Thus, the choice of $K$ does not appear to be a very critical consideration for these models unlike FF models.
+\\
 
@@
 This improvement in performance is not as pronounced as the CAE models for the case of speech mixtures.
-However, the experimentation serves as a proof of concept that exploring non-uniform auto-encoder architectures could lead to interesting and potentially powerful algorithms for several applications.
+However, the experimentation serves as a proof of concept that exploring non-uniform auto-encoder architectures could lead to interesting and potentially powerful algorithms for other datasets and applications.
+\\
 
@@
 Although these models are not as potent as CAE models for separation of speech mixtures, these extensions could lead to interesting non-uniform auto-encoder architectures for several applications.
+\\
 

### 2017-06-13 paper/paper.tex
Updated paper and pdf to latest version

@@
 \begin{abstract}
-We present a convolutiional auto-encoder that can act as a neural network equivalent to convolutive audio models.
+We present a convolutional auto-encoder that can act as a neural network equivalent to convolutive audio models.
 We demonstrate the ability of the network to learn optimal bases that resemble phoneme magnitude-spectra and also show how it can be used for supervised separation of speech mixtures.
@@
 Although the SAR for CAE models degrades slightly as compared to FF models, the significant improvement in SIR compensates for this loss.
+The convolutive speech models obtained by training the RCAE also outperform the FF auto-encoder models significantly, as seen by the median values and inter-quartile ranges.
+This improvement in performance is not as pronounced as the CAE models for the case of speech mixtures.
+However, the experimentation serves as a definite proof of concept that exploring non-uniform auto-encoder architectures could lead to interesting and potentially powerful algorithms for other datasets and applications.
+\\
+
 The performance of the models achieves a peak value for $K=80$.
@@
 Thus, the choice of $K$ does not appear to be a very critical consideration for these models unlike FF models.
-\\
-
-The convolutive speech models obtained by training the RCAE also outperform the FF auto-encoder models, as seen by the median values and inter-quartile ranges.
-This improvement in performance is not as pronounced as the CAE models for the case of speech mixtures.
-However, the experimentation serves as a proof of concept that exploring non-uniform auto-encoder architectures could lead to interesting and potentially powerful algorithms for other datasets and applications.
+At the same time, the variance in SDR seems to be dependent on $K$.
+We observe that the variance in SDR is considerably lower for higher values of $K$ ($K\geq50$) as seen by the separation results for both the convolutive models.
+In other words, auto-encoder based convolutive audio models produce superior results consistently for a high number of CNN filters.
 \\

### 2017-06-13 paper/paper.tex
I have made changes in paper.tex

@@
 Non-negative matrix factorization~(NMF) of magnitude-spectrograms has been a very popular method for modeling sources for supervised source separation applications~\cite{smaragdis2014static, virtanen2015compositional}.
-NMF factorizes a matrix of non-negative elements~$\mathbf{X}\in \mathbb{R}_{M\times N}^{\geq0}$ as a product of the basis matrix~$\mathbf{W}$ and the activation matrix~$\mathbf{H}$.
-The notation $\mathbb{R}_{M \times N}^{\geq0}$ represents the set of matrices of non-negative elements of size $M \times N$.
-In this factorization, the basis matrix $\mathbf{W} \in \mathbb{R}_{M \times r}^{\geq{0}}$, the activation matrix $\mathbf{H} \in \mathbb{R}_{r \times N}^{\geq{0}}$ and $r$ represents the rank of the decomposition.
-In the case of audio signals, we apply NMF on audio spectrograms.
-In the audio setting, the columns of $\mathbf{W}$ begin to act as representative basis vectors for the source.
+NMF factorizes a matrix of non-negative elements~$\mathbf{X}\in \mathbb{R}_{M\times N}^{\geq0}$ as a product of the basis matrix~$\mathbf{W}\in \mathbb{R}_{M \times r}^{\geq{0}}$ and the activation matrix~$\mathbf{H}\in \mathbb{R}_{r \times N}^{\geq{0}}$.
+The notation $\mathbb{R}_{M \times N}^{\geq0}$ represents the set of matrices of non-negative elements of size $M \times N$, and $r$ represents the rank of the decomposition.
+In the case of audio signals, NMF is applied on audio spectrograms, where the columns of $\mathbf{W}$ act as representative basis vectors for the source.
 The rows of $\mathbf{H}$ indicate the activity of these basis vectors in time.
-\textbf{C: I am not sure how smooth the transition to neural nets here.} As shown in~~\cite{smaragdis2017aneural}, the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network.
+\\
+
+As shown in~~\cite{smaragdis2017aneural}, the notion of non-negative audio modeling can be easily generalized by interpreting NMF as a neural network.
 We can interpret NMF as a non-negative auto-encoder in the following manner,
 \begin{align}
-    \text{$1^{\text{st}}$ layer:~}\mathbf{H} &= g(\mathbf{W^{\ddagger}} \cdot \mathbf{X}) \nonumber \\
-    \text{$2^{\text{nd}}$ layer:~}\mathbf{X} &= g(\mathbf{W} \cdot \mathbf{H})
+    \text{$1^{\text{st}}$ layer:~ (Encoder)~}\mathbf{H} &= g(\mathbf{W^{\ddagger}} \cdot \mathbf{X}) \nonumber \\
+    \text{$2^{\text{nd}}$ layer:~ (Decoder)~}\mathbf{X} &= g(\mathbf{W} \cdot \mathbf{H})
     \label{eq:nmfae}
@@
 Although non-negativity of the network-parameters (models) is not explicitly guaranteed in this formulation, applying a suitable sparsity constraint allows the network to learn suitable non-negative models.
-Additionally, this interpretation enables a pathway to propose multi-layer extensions by exploiting the wealth of available neural net architectures that could potentially lead to superior separation performance.
+Additionally, this interpretation enables a pathway to propose variants to this basic autoencoder stucture by exploiting the wealth of available neural net architectures that potentially lead to superior separation performance.
+The main alternative we explore in this paper is the convolutive architecture, which is based on convolutive NMF \cite{smaragdis2007convolutive}.
 \\
@@
 In this paper, we develop a neural network alternative to such convolutive audio models for supervised source separation.
-In doing so, we solve two fundamental problems associated with this task.
+\textbf{C:What does this following three sentences mean?}In doing so, we solve two fundamental problems associated with this task.
 The first step is to develop a suitable neural network architecture to learn convolutive audio models in an adaptive manner.
@@
 
-Developing a neural network alternative to conv-NMF also results in a significant advantage in terms of model re-usability.
-Several neural network architectures have been recently proposed for supervised source separation~\cite{grais2017single, venkataramani2017end}.
-These networks are trained to learn discriminative audio models by training them on mixture spectrograms.
-Thus, the networks learn a decomposition that works only for a specific source-interference pair.
-Developing an alternative to conv-NMF allows us to learn convolutive audio-models by training neural networks on clean source spectrograms.
-These models can be easily re-used irrespective of the interfering component.
-At the same time, it is also easy to propose extensions to the models by using different and exotic neural network layers.
+Several neural network architectures have been recently proposed for supervised source separation~\cite{grais2017single, venkataramani2017end}, which are discriminatively trained on audio mixtures.
+Although discriminative training of a source separation network results in good seperation performance, the network is restricted to work on a particular type of mixture.
+The convolutive architecture that we mainly discuss in this paper is generatively trained, and therefore not restricted to the types of mixtures that is used in training.
+By the virtue of flexibility of neural networks, we also propose a variant where recurrent neural network (RNN) is used.
 \\
@@
 
-\textbf{C: Impulse response concept is for linear systems.
-I am not too sure about using it in this context.} The decoder CNN can be interpreted as a finite impulse response~(FIR) filter.
-The encoder attempts to approximate the inverse of the decoder FIR filter as an FIR filter.
-From our knowledge of signal processing, it can be shown that the inverse filter of an FIR filter has an infinite impulse response~(IIR)~\cite{smith2011spectral}\footnote{\text{http://ccrma.stanford.edu/\~{}jos/fp/Inverse\_Filters.html}}.
-Thus, approximating the encoder as an FIR filter or a CNN will hinder the decoder from learning patterns that are invertible only by an IIR filter.
-Therefore, we explore using a Recurrent Neural Network~(RNN) to get an IIR encoder.
-\\
+Note that the encoder attempts to approximate the inverse of the decoder.
+From our knowledge of linear filtering in signal processing, we know that the inverse of a finite length filter is given by a recursive filter.
+~\cite{smith2011spectral}\footnote{\text{http://ccrma.stanford.edu/\~{}jos/fp/Inverse\_Filters.html}}.
+Following this analogy, in the next section we explore using a recurrent neural network in the encoder.
 
@@
 The potential gain of using a recurrent encoder over a finite length convolutive encoder is due to the fact that a recurrent filter in theory can capture arbitrarily long temporal dependencies.
-From a signal processing point of view, the motivation for going with a recurrent filter is that, the inverse of an finite length filter (the convolutive basis in the encoder) is given by a recurrent filter.
-We explore the idea, and obtain source separation results on the same dataset.
+From a signal processing point of view the motivation for going with a recurrent filter is that, the inverse of an finite length filter (the convolutive basis in the encoder) is given by a recurrent filter.
 \\

### 2017-06-13 paper/paper.tex
I have made few more changes on the paper.

@@
 For the evaluation, we generate $20$ such mixtures and compare the models for different parameter configurations.
-As a pre-processing step, we apply a $1024$ point short-time Fourier transform representation with a hop of $25\%$.
+As a pre-processing step, we apply a $1024$ point short-time Fourier transform representation with a hop size of $25\%$.
 The magnitude spectrogram is then given as an input to the network.
@@
 
-The neural networks are initialized using the Xavier initialization scheme~\cite{glorot2010understanding}.
 The networks are trained by applying a batch gradient descent training procedure and the parameters updated using the RMSProp algorithm~\cite{tieleman2012rmsprop}, with a learning rate and momentum of $0.001$ and $0.7$ respectively.
-\\
+The neural networks are initialized using the Xavier initialization scheme~\cite{glorot2010understanding}.\\
 
 The CNN filters are selected to be $512$-point tall and $8$-point wide.
-Thus, the convolutions are performed only along the time axis.
 The number of CNN filters also decides the number of components in the decomposition.
@@
 In order to maintain a uniform experimental setup, we apply the separation scheme described in~\ref{sec:ss} to all the models.
-We plot the results in terms of a violin plot that also indicates the median value (dotted line in the centre) and the corresponding inter-quartile range (outer dotted lines) for each $K$.
+We plot the results in terms of a violin plot that also indicates the median value (dotted line in the center) and the corresponding inter-quartile range (outer dotted lines) for each $K$.
 \\
@@
 
-\section{Conclusion}
+\section{Conclusions}
 \label{sec:conclusion}
@@
 One such extension considered in this paper is the use of auto-encoders formed by a cascade of recurrent and convolutive layers.
-Although these models are not as potent as CAE models for separation of speech mixtures, these extensions could lead to interesting non-uniform auto-encoder architectures for several applications.
+Although this extension have not outperformed the convolutive architecture, we have shown that this architecture outperforms the standard feedforward network.
 \\
